---
title: "ALA Data Providers"
description: |
  Analysis of data provider count summaries for selected taxa
author:
  - name: Martin Westgate 
    url: people/Westgate_Martin/index.html
# date: "15-12-2020"
categories: 
  - Taxa
  - Data Providers
output: 
  distill::distill_article:
    highlight: pygments
    toc: true
    toc_float: true
    toc_depth: 3
---

Task: look up largest data providers for the following groups:
* plants (kingdom plantae)
* birds (class aves)
* non-avian vertebrates
* invertebrates (non-chordate animals)

Note that not all of these are monophyletic groups. Hence, the best approach is to calculate:
* counts for kingdom animalia (A)
* counts for phylum chordata (B)
* counts for class aves (C)
* then sum to get the groups we want, i.e. birds (class aves) == C, non-avian vertebrates == B - C, invertebrates (non-chordate animals) == A - B


First download necessary packages

```{r, warning = FALSE, message = FALSE}
library(galah)
library(packcircles)
library(jsonlite)
library(crul)
library(ggplot2)
library(kableExtra)
```


Next find all taxa in Plantae, Animalia, Chordata and Aves

```{r}
taxa_list <- c("plantae", "animalia", "chordata", "aves")
id_list <- lapply(taxa_list, ala_taxa) # FIXME
names(id_list) <- taxa_list
```

Look up count

```{r}
# look up counts
dr_uid_list <- lapply(taxa_list, function(a){
  result <- ala_counts(id_list[[a]], 
                       breakdown = "data_resource_uid", 
                       limit = 1000)
  colnames(result) <- c("dr_uid", a)
  return(result)
})
names(dr_uid_list) <- taxa_list
```

Merge the counts data into a single data frame

```{r}
# merge into a single data.frame
dr_uid <- merge(dr_uid_list$plantae, dr_uid_list$animalia, 
                by = "dr_uid", 
                all = TRUE)
dr_uid <- merge(dr_uid, dr_uid_list$chordata, 
                by = "dr_uid", 
                all = TRUE)
dr_uid <- merge(dr_uid, dr_uid_list$aves, 
                by = "dr_uid", 
                all = TRUE)
for(i in 2:5){dr_uid[is.na(dr_uid[, i]), i] <- 0}
```

Next we create new columns that contain the sum of counts

```{r}
# create new summed columns
dr_uid$invertebrates <- dr_uid$animalia - dr_uid$chordata
dr_uid$nonavian_vertebrates <- dr_uid$chordata - dr_uid$aves
dr_uid <- dr_uid[, !(colnames(dr_uid) %in% c("animalia", "chordata"))]
dr_uid <- dr_uid[, c(1, 2, 4, 5, 3)]

# reduce total number of entries
dr_uid <- dr_uid[apply(dr_uid[, -1], 1, sum) > 10000, ]
```

Now extract the taxon information and generate a best-possible-guess of the data provider

```{r}
# get taxon info
uid_list <- lapply(dr_uid$dr_uid, lookup_resource_uid)
# if the above doesn't work, try in a loop to check for bugs
# uid_list <- as.list(rep(NA, nrow(dr_uid)))
# for(i in c(53:length(uid_list))){
#   uid_list[[i]] <- lookup_resource_uid(dr_uid$name[i])
# }

# generate the best possible guess of data provider
uid_df <- do.call(rbind, uid_list)[, c(1, 3, 4, 2)]
```

We also need to add a specific call to add CSIRO National Fish COllection data which are not listed as OZCAM data for some reason

```{r}
# CSIRO National Fish Collection not listed as OZCAM for some reason
ozcam_text <- "OZCAM (Online Zoological Collections of Australian Museums) Provider"
uid_df$provider[which(grepl("OZCAM", uid_df$name))] <- ozcam_text
uid_df$provider[which(grepl("ANIC", uid_df$name))[2]] <- ozcam_text
uid_df$provider[which(uid_df$provider == "Australian National Insect Collection, CSIRO")] <- ozcam_text
```

# Create labels

Now we want to create readable labels with useful information about providers

```{r}
# generate a label with the most useful information about that provider
uid_df$label <- apply(uid_df[, c(2:4)], 1, function(a){
  if(all(is.na(a))){
    NA
  }else{
    a[min(which(!is.na(a)))]
  }
})

# CSIRO is listed using many different labels - standardize
uid_df$label[which(
  (grepl("Commonwealth Scientific and Industrial Research Organisation", uid_df$label) |
  grepl("CSIRO", uid_df$label)) &
  !grepl("OZCAM", uid_df$provider)
)] <- "CSIRO"

# combine herbaria
uid_df$label[
  uid_df$label %in% c("New Zealand Virtual Herbarium", "Australia's Virtual Herbarium")
] <- "Australasian Virtual Herbarium"

# add this content back to the count dataset
dr_uid$provider <- uid_df$label
```

Some providers have multiple uids, so we now need to group repeat observations

```{r}
# Because some providers have multiple uids, we now need to group repeat obs
provider_list <- split(dr_uid, dr_uid$provider)
count_df <- as.data.frame(do.call(rbind, lapply(provider_list, function(a){
  c(
    provider = a$provider[1],
    apply(a[, 2:5], 2, function(b){sum(b, na.rm = TRUE)})
  )
})))
rownames(count_df) <- NULL
count_list <- lapply(
  colnames(count_df)[2:5],
  function(a){data.frame(provider = count_df$provider, count = as.numeric(count_df[, a]))})
names(count_list) <- colnames(count_df)[2:5]
count_list <- lapply(count_list, function(a){a[order(a$count, decreasing = TRUE), ]})
count_list <- lapply(count_list, function(a){a[1:10, ]})

# convert to df
count_df <- as.data.frame(do.call(rbind, count_list))
rownames(count_df) <- NULL
count_df$taxon <- rep(names(count_list), each = 10)
```

Now we will manually write better labels. This is messy code, but it works!

```{r}
# manually write better labels (messy code, but works!)
count_df$label <- NA

count_df$label[grepl("Office of Environment and Heritage", count_df$provider)] <- "Office of Environment\n& Heritage (NSW)"
count_df$label[grepl("Victorian Department of Environment", count_df$provider)] <- "DELWP (VIC)"
count_df$label[grepl("South Australia, Department for Environment", count_df$provider)] <- "Dept. Env.\n& Water (SA)"
count_df$label[grepl("Northern Territory Department of Environment", count_df$provider)] <- "Dept. Env. & Natural\nRes. (NT)"
count_df$label[grepl("Tasmanian Department of Primary Industries", count_df$provider)] <- "DPIPWE (TAS)"
count_df$label[grepl("OZCAM", count_df$provider)] <- "Museums & Collections"
count_df$label[grepl("eBird", count_df$provider)] <- "eBird"
count_df$label[grepl("BirdLife", count_df$provider)] <- "BirdLife Australia"

count_df$label[count_df$provider == "Australasian Virtual Herbarium"] <- "Australasian Virtual Herbarium"
count_df$label[count_df$provider == "Ocean Biogeographic Information System"] <- "OBIS"
count_df$label[count_df$provider == "WildNet - Queensland Wildlife Data"] <- "WildNet (QLD)"
count_df$label[count_df$provider == "iNaturalist Australia"] <- "iNaturalist"
count_df$label[count_df$provider == "NatureMap"] <- "NatureMap"
count_df$label[count_df$provider == "Australian Antarctic Data Centre"] <- "AADC"
count_df$label[count_df$provider == "Murray-Darling Basin Authority"] <- "MDBC"
count_df$label[count_df$provider == "CSIRO"] <- "CSIRO"
count_df$label[count_df$provider == "Questagame"] <- "Earth Guardians"
count_df$label[count_df$provider == "BowerBird"] <- "BowerBird"
count_df$label[count_df$provider == "Reef Life Survey"] <- "Reef Life Survey"
count_df$label[count_df$provider == "NSW Bird Atlassers"] <- "NSW Bird\nAtlassers"
count_df$label[count_df$provider == "Canberra Ornithologists Group"] <- "Canberra Ornithologists Group"

saveRDS(count_df, "labelled_count_df_highTaxa_v2.rds")
# count_df <- readRDS("./data/labelled_count_list_highTaxa.rds")
```



# Plotting

## Plot building functions

First, we create some functions to do some more complex stuff that we will repeat for several plots.

Make the grey circle polygon of specified sizes

```{r}
# functions for creating circular polygons of specified size
make_circle <- function(
  n = 100, # number of points, equally spaced around the edge of a circle
  k,  # scaling value (radius) - larger for bigger circles. defaults to 1
  offset = c(0, 0)
){
  if(missing(k)){k <- 1}
  # get coordinates
  theta <- (2 * (pi/(n-1)) * seq_len((n-1)) ) # -alpha
  values <- data.frame(
  	x = k * cos(theta),
  	y = k * sin(theta))
  rownames(values) <- NULL
  values$x <- values$x + offset[[1]]
  values$y <- values$y + offset[[2]]
  # return(values)
  return(values[c(seq_len(n-1), 1), ])
}
# note the result has nrow == n, but n-1 unique rows
```

Make circles, but in a loop

```{r}
# run make circle in a loop
make_circles <- function(
  df, # should be returned by packcircles::circleProgressiveLayout
  gap = 0
){
  df_list <- split(df, seq_len(nrow(df)))
  result <- as.data.frame(do.call(rbind,
    lapply(df_list, function(a){
      make_circle(k = a$radius - gap, offset = c(a$x, a$y))
    })
  ))
  result$group <- rep(seq_along(df_list), each = 100)
  return(result)
}
```

Build a circle-packing plot

```{r}
# function to build a circle-packing plot
# this is a bit over-engineered for this purpose but works ok
taxon_plot_data <- function(
  labels, # vector of taxon labels
  counts, # vector of counts
  color_scale = "viridis",
  log_scale = TRUE, # logical - should the result be log + 1 -scaled?
  gap = 0 # how much should the radius be reduced to decrease circle overlap?
){
  if(log_scale){
    taxon_counts <- log(counts + 1)
  }else{
    taxon_counts <- counts
  }
  circle_df <- packcircles::circleProgressiveLayout(taxon_counts)
  circle_plot_df <- make_circles(circle_df, gap = gap)
  circle_plot_df$label <- factor(circle_plot_df$group,
    levels = unique(circle_plot_df$group),
    labels = labels)
  circle_plot_df$count <- rep(counts, each = 100)
  circle_plot_df$plotly_text <- paste0(
    "<b>",
    circle_plot_df$label,
    "</b><br>n = ",
    formatC(circle_plot_df$count, big.mark = ","))
  circle_plot_df$color <- rep(
    do.call(color_scale, args = list(
      n = length(labels),
      direction = -1)),
    each = 100)
  return(circle_plot_df)
}
```

Query the data provider API

```{r}
# query the data provider API
lookup_resource_uid <- function(id){
  url <- paste0("https://collections.ala.org.au/ws/dataResource/", id)
  cli <- HttpClient$new(url)
  cli <- cli$get()
  text <- cli$parse("UTF-8")
  if(nchar(text) < 10000){
    result <- fromJSON(text, flatten = TRUE)
    out <- data.frame(
      uid = {if(is.null(result$uid)){NA}else{result$uid}},
      name = {if(is.null(result$name)){NA}else{result$name}},
      provider = {if(is.null(result$provider$name)){NA}else{result$provider$name}},
      institution = {if(is.null(result$institution$name)){NA}else{result$institution$name}}
    )
  }else{
    out <- data.frame(uid = NA, name = NA, provider = NA, institution = NA)
  }
  return(out)
}
```

Set colours

```{r}
# plot code
ALA_colors <- list(
  flamingo = "#E06E53", # primary
  rust = "#B8573E",
  grey = "#667073",
  concrete = "#EEECEA", # secondary (monochrome)
  silver = "#9E9E9F",
  charcoal = "#222322",
  honey = "#FFC557", # extended
  pale_moss = "#B7CD96",
  seafoam = "#6BDAD5",
  ocean = "#003A70",
  # lavender = "#A19B2",
  plum = "#691C32"
)
scales::show_col(unlist(ALA_colors))
```

Make the plot!

```{r}
plot_providers <- function(data_df, labels_df, label_split = 2000, color = "#E06E53"){
  ggplot() +
    geom_polygon(
      data = data_df,
      mapping = aes(x = x, y = y, group = group, alpha = -group),
      fill = color
    ) +
    # add centred text for large circles
    geom_text(
      data = labels_df[labels_df$count >= label_split, ],
      mapping = aes(x = x, y = y, label = label),
      hjust = 0.5,
      color = "white",
      size = 3
    ) +
    # add offset labels for small circles
    geom_text(
      data = labels_df[labels_df$count < label_split, ],
      mapping = aes(x = x_max, y = y, label = label),
      color = "#667073",
      hjust = 0,
      nudge_x = 50,
      size = 3
    ) +
    # optional code for adding counts
    # geom_text(
    #   data = labels_df[labels_df$x > x_split, ],
    #   mapping = aes(x = x_max, y = y, label = count),
    #   color = "#667073",
    #   hjust = 0,
    #   nudge_x = 50,
    #   nudge_y = y_offset,
    #   size = 2
    # ) +
    # geom_text(
    #   data = labels_df[labels_df$x <= x_split, ],
    #   mapping = aes(x = x, y = y, label = count),
    #   color = "white",
    #   hjust = 0.5,
    #   nudge_y = y_offset,
    #   size = 2
    # ) +
    coord_fixed() +
    # optional content for changing the color scheme
    # scale_fill_viridis(option = "magma", begin = 0.1, end = 0.9, direction = -1) +
    # scale_fill_gradient(low = "#B8573E", high = "#E06E53") +
    scale_alpha(range = c(0.5, 1)) +
    theme_void() +
    theme(
      legend.position = "none",
      plot.title = element_text(hjust = 0.5)
    )
}
```


## Build plot

Now we can generate plot data

```{r}
# generate plot data
library(viridis)
plot_list <- lapply(split(count_df, count_df$taxon)[c(4, 2, 3, 1)], function(a){
  result <- taxon_plot_data(a$label, a$count, log_scale = FALSE)
  result$count_label <- formatC(result$count, format = "f", digits = 0, big.mark = ",")
  return(result)
})

label_list <- lapply(plot_list, function(a){
  do.call(rbind, lapply(
    split(a, a$group),
    function(b){
      data.frame(
        x = mean(b$x),
        x_max = max(b$x),
        y = mean(b$y),
        count = b$count[1], # b$count_label[1],
        label = b$label[1])
    })
  )
})
```

Create plots using the `plot_providers` function for:

Plants
```{r}
# plot this
# plantae
plot_providers(plot_list$plantae, label_list$plantae,
  label_split = 900000,
  color = ALA_colors[["pale_moss"]]) +
  ggtitle("Providers of plant data (kingdom Plantae) to ALA") +
  lims(x = c(-3000, 5000))
# ggsave("./plots/final_provider_plot_plantae.pdf")
```


Invertebrates
```{r}
# invertebrates
plot_providers(plot_list$invertebrates, label_list$invertebrates[1:6, ],
  label_split = 110000,
  color = ALA_colors[["seafoam"]]
) +
  geom_text(
    data = label_list$invertebrates[7:9, ],
    mapping = aes(x = x_max, y = y, label = label),
    color = "#667073",
    hjust = 0,
    nudge_x = 50,
    nudge_y = -50,
    size = 2.5
  ) +
  geom_text(
    data = label_list$invertebrates[10, ],
    mapping = aes(x = x, y = y, label = label),
    color = "#667073",
    hjust = 0.5,
    # nudge_x = 100,
    nudge_y = 220,
    size = 3
  ) +
  ggtitle("Providers of invertebrate data to ALA") +
  lims(x = c(-2300, 2000))
# ggsave("./plots/final_provider_plot_invertebrates.pdf")
```

Non-avian vertebrates
```{r}
# non-avian vertebrates
plot_providers(plot_list$nonavian_vertebrates, label_list$nonavian_vertebrates[1:8, ],
  label_split = 500000,
  color = ALA_colors[["flamingo"]]
) +
  geom_text(
    data = label_list$nonavian_vertebrates[9:10, ],
    mapping = aes(x = x_max, y = y, label = label),
    color = "#667073",
    hjust = 0,
    nudge_x = 20,
    nudge_y = -150,
    size = 3
  ) +
  ggtitle("Providers of non-avian vertebrate data to ALA") +
  lims(x = c(-2000, 3000))
# ggsave("./plots/final_provider_plot_nonavian_vertebrates.pdf")
```

Aves
```{r}
# aves
plot_providers(plot_list$aves, label_list$aves[-3, ],
  label_split = 3000000,
  color = ALA_colors[["honey"]]
) +
  geom_text(
    data = label_list$aves[3, ],
    mapping = aes(x = x_max, y = y, label = label),
    color = "#667073",
    hjust = 0,
    nudge_x = 50,
    nudge_y = -500,
    size = 3
  ) +
  ggtitle("Providers of bird data to ALA") +
  lims(x = c(-5000, 8000))
# ggsave("./plots/final_provider_plot_aves.pdf")
```


