


```{r}
library(galah)
library(here)
library(rmapshaper)
library(tidyverse)
library(sf)
library(ozmaps)
library(tidyr)
library(ggpointdensity)
library(glue)
install.packages("ggtext")
library(ggtext)
library(prismatic)
library(magrittr)
```

You can put the CAPAD folder in the projects/dingo/data folder and this will run

here()
here::here("/Users/smi9e9/Documents/tutorials")

```{r}
capad <- st_read(here("/Users",
                      "smi9e9",
                      "Documents",
                      "tutorials",                     
                      "CAPAD",
                      "CAPAD2020_terrestrial.shp"),
                 quiet = TRUE) |>
  ms_simplify(keep = 0.1)
```

```{r}
summary(capad)
view(capad_counts)
```

```{r}
capad <- capad |>
  st_transform(crs = st_crs("WGS84")) |>
  st_make_valid()
```


```{r}
install.packages("prismatic")
```


```{r}
terrain10 <- terrain.colors(10)

terrain10
#>  [1] "#00A600" "#2DB600" "#63C600" "#A0D600" "#E6E600" "#E8C32E" "#EBB25E"
#>  [8] "#EDB48E" "#F0C9C0" "#F2F2F2"

terrain_color <- color(terrain10)

terrain_color
#> <colors>
#> #00A600FF #2DB600FF #63C600FF #A0D600FF #E6E600FF #E8C32EFF #EBB25EFF #EDB48EFF #F0C9C0FF #F2F2F2FF

plot(terrain_color)
```
```{r}
color(rainbow(64))
```


```{r}
galah_config(email = "SMI9E9@csiro.au", verbose = FALSE)
```

```{r}
dingoocc <- galah_call() |>
  galah_identify("canis familiaris") |>
  galah_filter(raw_vernacularName == "Dingo") |>
  galah_apply_profile(ALA) |>
  galah_group_by(dataResourceName, cl22) |>
  atlas_occurrences()
```

This is a simpler way to prepare the points for plotting if you want to plot using `geom_sf()`**

```{r}
# convert occurrence points to `sf` object, drop records without lat or lon coords, add projection
dingo_sf <- dingoocc |> 
  drop_na(decimalLongitude, decimalLatitude) |>
  st_as_sf(coords = c("decimalLongitude", "decimalLatitude"), 
           crs = st_crs("WGS84"))
```


From this point, you should have everything you need to make the map**
- I think it would be nice to have a white map of australia underneath the CAPAD layer. You can get this from the {ozmaps} package
- Then you can plot the dingo observations points on the map. Using `alpha` could help make it easier to see which points are in the national parks
- Decide whether looking at all of Australia is too difficult to see the points and the national parks. Maybe it would be better to filter the map to only show one state at a time? I think you should be able to do that with the code that is here :)


```{r}

capad_counts <- capad |> 
  mutate(dingo_count = pmap(.l = list(x = capad$geometry),
                           .f = function(x) {
                             lengths(st_intersects(x, dingo_sf))
                             }))
```

```{r}
dingoNT <- capad_counts |> select(STATE, dingo_count, NAME) |>
  filter(STATE == "NT")
```

```{r}
hist(as.numeric(capad_counts$dingo_count))
```

```{r}
dingoNT |>
  arrange(desc(dingo_count)) 
#still need to get this to work

view(dingoNT)
```


```{r}
dingo_palette <- list("AustralianProtectedAreas" = "#00A600FF")

                      #"Dingoobservations" = " viridis::scale_color_viridis")
```

```{r}
galah_config(email = "SMI9E9@csiro.au")

dingo_records <- galah_call() |>
  galah_identify("canis familiaris") |>
  galah_filter(raw_vernacularName == "Dingo")|>
  # galah_filter(cl22 == "NT")
  atlas_occurrences()
```


```{r}
# Tidy
dingo_records <- dingo_records |>
  drop_na() |>                      # drop NAs
  filter(decimalLongitude < 155,    # filter records to only mainland/tasmania
         decimalLongitude > 110, 
         decimalLatitude > -45,
         decimalLatitude < -10)
```

```{r}
capad <- st_read(here("/Users",
                      "smi9e9",
                      "Documents",
                      "tutorials",                     
                      "Transport",
                      "Roads_Principle_Aust.shp"),
                 quiet = TRUE) |>
  ms_simplify(keep = 0.1)
```
```{r}
roads <- capad |>
  st_transform(crs = st_crs("WGS84")) |>
  st_make_valid()
```

```{r}
view(roads)
```

```{r}
Dog_Fence <- st_read(here("/Users",
                      "smi9e9",
                      "Documents",
                      "tutorials",                     
                      "Dog Fence",
                      "Dog_Fence.shp"),
                 quiet = TRUE) |>
  ms_simplify(keep = 0.1)
```


```{r}
Dog_Fence <- Dog_Fence |>
  st_transform(crs = st_crs("WGS84")) |>
  st_make_valid()
```

I'm also unsure of how to get the records just in the NT, I have tried the group by function and filter but doesnt seem to work f

This is the point density one which I prefer more. 

```{r}
NT_plot <- ggplot() +
  geom_sf(data = ozmap_states, 
          fill = "beige", colour = "black", linewidth = 0.3) +
  geom_sf(data = capad, fill = "#00A600FF", colour = "black", linewidth = 0.25, alpha = 0.3, linetype = "dashed") +
  geom_pointdensity(data = dingo_records, size = 1, alpha = 0.5,
                    mapping = aes(x = decimalLongitude,
                                  y = decimalLatitude)) +
    viridis::scale_color_viridis(option = "B", 
                               begin = 0.1,
                               guide_colorbar(
                                              title = ("Number of \nOverlapping observations"), 
                                              alpha = "none")) +
  coord_sf(xlim = c(128, 140), ylim = c(-10, -26)) + # filter to mainland australia+
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0), 
        panel.background = element_rect(fill = "white", size = 2), 
        panel.grid.major = element_line(colour = NA),
        legend.position = "bottom",
        plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
        axis.ticks = element_line(colour = "grey70", size = 0.2),
        axis.title.y = element_text(size = 12),
        axis.title.x = element_text(size = 12)) +
        labs(title = "Dingo obsverations in 
             Australian Protected Areas in the Northern Territory (NT)", x = "Longitude", y = "Latitude") 
```

```{r}
library(ggnewscale)
```




```{r}
NT_plot
```
```{r}
labelled_plot <- NT_plot +
  ggnewscale::new_scale_color() +
  scale_colour_manual(values = dingo_palette) +
  labs(title = glue("in <span style='color:{dingo_palette$AustralianProtectedAreas}'>**AustralaianProtectedAreas**</span> <br> in the Northern Territory")) +
  theme(plot.subtitle = element_markdown())
```

```{r}
labelled_plot
```


```{r}
table(dingoNT$dingo_count, dingoNT$NAME)
```


```{r}
labelled_plot
```

```{r}
SA_plot <- ggplot() +
  geom_sf(data = ozmap_states, 
          fill = "beige", colour = "black", linewidth = 0.3) +
  geom_sf(data = capad, fill = "#00A600FF", colour = "black", linewidth = 0.25, alpha = 0.3, linetype = "dashed") +
  geom_sf(data = Dog_Fence, color = "#0800FFFF", linewidth = 2) +
  geom_pointdensity(data = dingo_records, size = 1, alpha = 0.5,
                    mapping = aes(x = decimalLongitude,
                                  y = decimalLatitude)) +
    viridis::scale_color_viridis(option = "F", 
                               begin = 0.1,
                               guide_colorbar(direction = "horizontal", label.position = "left",
                                              title = ("Number of \nOverlapping observations"), 
                                              alpha = "none")) +
  coord_sf(xlim = c(128, 145), ylim = c(-40, -26)) + # filter to mainland australia
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0), 
        panel.background = element_rect(fill = "white", size = 2), 
        panel.grid.major = element_line(colour = NA),
        plot.title = element_text(face = "bold", size = 16, hjust = 0.5, line = 1),
        axis.ticks = element_line(colour = "grey70", size = 0.2),
        axis.title.y = element_text(size = 12),
        axis.title.x = element_text(size = 12)) +
        labs(title = 
        "Dingo obsverations in 
        Australian Protected Areas in South Australia (SA)", x = "Longitude", y = "Latitude")
```

```{r}
SA_plot
```

