---
title: "Good Practices in R scripts"
author: "ALA and EcoCommons | June 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Document created for a joint decision on best practices to be adopted on R scripts within ALA and EcoCommons, 
aiming consistency and a common vocabulary. While scripts usually involves personal styles, it would be nice 
to follow some general rules. Please add comments and pieces of information on what you usually do. Right now, 
this is just a draft with some practices I follow and some that I found on the internet - Jess

***

> ### **Useful sites:**

* [Google's R Style Guide](https://google.github.io/styleguide/Rguide.html)
* [Hadley Wickham’s Style Guide](http://adv-r.had.co.nz/Style.html)
* [Environmental Computing UNSW](http://environmentalcomputing.net/good-practice-for-writing-scripts/)


> ### **General script format**

* Start with a title, author details, a brief description of the scripts purpose
  and the data being used and copyright and legal stuff
  
* In the begging of the script, include commands to set working directory and 
  load necessary libraries
  
* Finish scripts with writeLines('Done')

EXAMPLE
```{r}
############################################################
###           BCCVL script to define parameters          ###
############################################################
##
## Author details: BCCVL, Contact details: emailadress@bccvl.au
## Copyright statement: This script is the product of BCCVL etc.
## Date : June 2021
## Script and data info:
## In this script you will create functions to set up parameters that
## will allow you to run multiple algorithms for species distribution
## modelling (SDMs)

#setwd("C:/Users/JessicaFenker/projectdirectory")
#library(biomod2)
#library(dismo)

#my.data <- rjson::fromJSON (file="mydata.json")

#...

#writeLines('Done')
```


> ### **Notation and Naming**

* Avoid weird separators (as backslash, symbols, space and punctuation marks)
  when naming your files
  
* Give your files short and meaningful names

* File names should be unique - avoid use the same name for other documents


> ### **Comments**

* Descriptive file names can help to avoid unnecessary comments

* Comments should not state the obvious

* Start comments with #,  one space and Capital letter. Short codes can be placed
  after the code, with two spaces, #, and then one space
  
* Functions need special comments, with one sentence description of the function,
  a list of arguments with a description (including data type) and description
  of the return value
  
EXAMPLE
```{r}
CalculateStandardError <- function (x){
  # Computes the sample standard error
  #
  # Arguments:
  #  x: Vector whose standard error is to be calculated. x must have length greater than one,
  #     with no missingn values.
  #
  # Return:
  #  The standard error of x
  se<-sd(x)/sqrt(length(x))
  return(se)
}
```
  
  
 
> ### **Sythax**

* Always use <- when assigning names to objects and avoid using = for assignment
*Even though this distinction doesn’t matter for the majority of the time, it is
 a good  habit to use <- as this can be used anywhere, whereas the operator = is
 only allowed at the top level. In addition = closely resembles ==, which is the 
 logical operator for equals to*
 
* Avoid long line length (~80 characters)

* Place spaces around all binary operators (=, +, -, <-, ==, ! = )

* Always put a space after a comma and never before

* Curley braces opening should never go on its own line and should always be
  followed by a new line. A closing curly brace should always go on its own line,
  unless followed by else, which should be contained within outward facing curly
  braces >}else{ . Indent the code within curly braces
  
* Always name VARIABLES with lower case and with words separated by underscore?? 
  (what do you think?) The use of nouns is a good practice; avoid single letters
  
* Name FUNCTIONS with initial capital letters and no dots. Use verbs
 *OR*
* Private FUNCTIONS should begin with a dot 

> ### **Pypes**

* Use explicit returns - do not rely on R's feature

* Explicitly qualify namespaces for external functions

EXAMPLE
```{r}
#purrr::map()
```


***

#### *Is there something missing?? Please contribute :)*
 *(and feel free to correct my grammar mistakes)*

***
