---
title: "Proportion of dingo records INSIDE vs OUTSIDE of protected areas"
author: "Amos Smith"
date: "2023-01-31"
output: html_document
---

# Get data

```{r}
# load packages
library(galah)
library(tidyverse)
library(ozmaps)
library(sf)
library(rmapshaper)
library(patchwork)
library(here)
```

Read shapefile

```{r}
capad <- st_read(here("projects",
                      "dingo",
                      "data",
                      "CAPAD",
                      "CAPAD2020_terrestrial.shp"),
                 quiet = TRUE) |>
  ms_simplify(keep = 0.1)
```

Download dingo records

```{r}
dingoocc <- galah_call() |>
  galah_identify("canis familiaris") |>
  galah_filter(raw_vernacularName == "Dingo") |>
  galah_apply_profile(ALA) |>
  galah_group_by(dataResourceName, cl22) |>
  atlas_occurrences()
```

# Wrangle data

Convert data to correct format

```{r}
# convert occurrences to sf object
dingo_sf <- dingoocc |>
  drop_na() |>
  st_as_sf(coords = c("decimalLongitude", "decimalLatitude"), 
  crs = st_crs("WGS84"))
```

Filter shapefile to only protected areas in the NT

```{r}
# filter
dingo_state <- capad |> 
  dplyr::select(STATE, NAME) |>
  dplyr::filter(STATE == "NT")
```

# Get counts

```{r}
# Count number of observations in each protected area
capad_counts <- dingo_state |> 
  mutate(dingo_count = pmap(.l = list(x = dingo_state$geometry),
                            .f = function(x) {
                             lengths(st_intersects(x, dingo_sf))
                             }))



# total in protected areas
counts_list <- capad_counts %>%
  pull(dingo_count)

counts_by_area <- tibble(name = capad_counts$NAME,
                         count = do.call(rbind, x))
```

# Results

Top 10 Protected areas with most records

```{r}
counts_by_area %>% 
  slice_max(count, n = 10)
```
Number of protected areas with no records

```{r}
counts_by_area %>% 
  filter(count == 0) %>%
  count()
```

Number of dingo records in protected areas

```{r}
total_state <- counts_by_area %>%
  summarise(total = sum(count))
```

Number of dingo records overall in the state

```{r}
total_overall <- galah_call() |>
  galah_identify("canis familiaris") |>
  galah_filter(raw_vernacularName == "Dingo",
               cl22 == "Northern Territory") |>
  galah_apply_profile(ALA) |>
  atlas_counts()
```

Proportion of records in Protected areas vs outside of Protected areas

```{r}
(total_state/total_overall) * 100
```



