---
title: "An exploration of dingo observations in the ALA"
author: "Amos Smith"
date: "2022-01-27"
output: 
  html_document:
    css: style.css
    highlight: tango
---

Pest or Native species? This is an ongoing debate for our nationally treasured dingo with each state having different legislation whether it is protected or not. The dingo is under threat from human wildlife conflict around agriculture, habitat loss and breeding with domestic dogs losing its genetic 'purity'. From this, they are listed as Vulnerable to extinction under the International Union of Nature Conservation’s Red List of Threatened Species.

Dingoes have been heavily culled across Australia from broad - scale baiting and shooting programs because of the impact they have on the agricultural industry and are only found to be protected within reserves. Its estimated the impact they have on the industry cause 1900 sheep deaths annually. 

From an ecological perspective, dingoes are crucial to regulating ecological cascades as an apex predator decreasing the abundance and impacts herbivores have by trampling on native vegetation. Also controls invasive mesopredator numbers like the red fox (Vulpus vulpes) and from this the loss of dingoes, has lead to widespread losses of small and medium -  sized native mammals and plant biomass. 

In South Australia, the dingo fence is the longest fence in the world, purposed to keep dingoes on one side and exterminated on the other because of the impacts it has on the sheep industry. This turned to be a natural experiment showing the removal of apex predators transformed an entire ecosystem and could be visually seen from space. On the side where dingoes remain present, there is more mammals, natural vegetation cover from less kangaroos and overall greater biodiversity. On the other side where dingoes have not been present for over 80 years in the region, there is more red foxes, feral cats and kangaroos which has denuded the natural vegetation. Too visually see this from space see here:  https://youtu.be/ivmQk5EsScw. 

Investigating species within the Atlas and understanding the factors that effect the data collection of a species can be key to better data collection of a species for future conservation, therefore we are going to investigate the dingo. 


First lets install the required packages.

```{r}
library(galah)
library(tidyverse)
library(ozmaps)
library(sf)
library(ggplot2)
library(patchwork)
library(lubridate)
library(here)
library(rmapshaper)
library(ggpointdensity)
library(glue)
library(ggtext)

```

First we should know how many counts are within the Atlas. 

```{r}
count <- galah_call()|>
  galah_identify("canis familiaris")|>
  galah_filter(raw_vernacularName == "Dingo")|>
  galah_apply_profile(ALA)|>
  atlas_counts()
count
```


Register email to the galah package. 

```{r}
galah_config(email = "SMI9E9@csiro.au")
```


#Which states have the highest counts of dingoes? 

We can now acquire the counts for the dingo by each state. Within the Atlas dingoes fall under the class canis familiaris which is also wild and domesticated dogs, therefore we had to filter records that were only recorded as dingoes. This can potentially mean that we have missing observations because data collectors did not specify that it was a dingo even if it was. We also filter our records that have been verified by the ALA so we aren't getting unreliable observations. 

```{r}
dingo_counts <- galah_call()|>
  galah_identify("canis familiaris")|>
  galah_filter(raw_vernacularName == "Dingo")|>
  galah_group_by(cl22)|>
  galah_apply_profile(ALA)|>
  atlas_counts()
```

Arrange the counts in descending order. 

```{r}
dingo_counts |>
  arrange(desc(count))
```
From this information it is obvious that the Northern Territory and South Australia hold the highest counts where in some states have very minimal observations. If we want to visually see this on a chloropleth map we will run the following code. 

This gives us our base map of Australia. 

```{r}
aus <- ozmap_data(data = "states")
```

Join the counts by state to our base map. 

```{r}
state_counts <- aus %>%
  full_join(dingo_counts, by = c("NAME" = "cl22"))
```

Log the counts by 10 to create a smoother distribution of the data by states. This will categorize it into counts by 10,100 and 1000.

```{r}
logged_counts <- state_counts |>
  mutate(log = log10(count))|>
  replace_na(list(log = 0))
```

We can plot our chloropleth map. 

```{r}
Chloropleth_map <- ggplot() +
  geom_sf(data = logged_counts,
          mapping = aes(fill = log,)) +
   scale_fill_distiller(palette = "OrRd",
                        direction = 1,
                        labels = c("10", "100", "1000"),
                        guide = guide_colorsteps(direction = "horizontal",
                        title = "No. of Dingo Observations",
                        title.position = "top")) +
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0), 
        panel.background = element_rect(fill = "white", size = 2), 
        panel.grid.major = element_line(colour = NA),
        plot.title = element_text(face = "bold", size = 16, hjust = 0.5, line = 1),
        axis.ticks = element_line(colour = "grey70", size = 0.2),
        axis.title.y = element_text(size = 12, vjust = 3),
        axis.title.x = element_text(size = 12, vjust = -1)) +
   theme(legend.position = "bottom", legend.key.width = unit(12, 'mm')) +
    labs(x = "Longitude", y = "Latitude")

Chloropleth_map
```
#Who are our main data providers? 

Why is it that majority of dingo observations are only occurring in the Northern Territory in comparison to Western Australia? We know that dingoes do occur in Western Australia but are not recorded into the Atlas. To understand who is currently recording them, we decided to break down our top data providers of the dingo into the atlas. This may give an idea of the absence of data of who could be recording dingoes but isn't. 


```{r}
dingo_dataproviders_count <- galah_call()|>
  galah_identify("canis familiaris")|>
  galah_filter(raw_vernacularName == "Dingo")|>
  galah_group_by(dataResourceName)|>
  galah_apply_profile(ALA)|>
  atlas_counts()
```

Arrange it in descending order. 

Here we can see that it is mainly state organisations recording the data, with 0 public/citizen recordings. With 0 public records it limits the ability to span our knowledge of its distribution from incidental records that can be in areas where state organisations aren't surveying for dingoes. For instance in WA, there is little observations on dingoes, citizen science can then become very valuable where its filling in those gaps that state organisations aren't able to cover for a certain species. Yes, it will have some bias to the data but in the long run we are still getting a better idea of where they are occurring. 

```{r}
dingo_dataproviders_count |>
  arrange(desc(count))
```


 If we want to visually see this on a map we can run the following code. 



This will obtain the dingo occurrences rather than counts which is useful for mapping because they obtain longitudinal and latitudinal values. 

```{r}
dingoocc_dataproviders <- galah_call()|>
  galah_identify("canis familiaris")|>
  galah_filter(raw_vernacularName == "Dingo")|>
  galah_group_by(dataResourceName)|>
  galah_apply_profile(ALA)|>
  atlas_occurrences()
```


To view the main data providers, we'll filter to only providers that have provided more than 5 observations of dingoes

```{r}
counts_fitered <- dingo_dataproviders_count |>
  filter(count > 5)
```

This filters our data frame with occurrences so that only the observations from each data provider with greater than 5 observations (in our `counts_filtered` data frame) are returned
```{r}
points_fitered <- dingoocc_dataproviders |>
  filter(dataResourceName %in% counts_fitered$dataResourceName)
```

Now we can plot

From this point, we have found our main data providers for the Atlas, which states in Australia have the highest counts and majority of records are recorded during winter. There are only 4 main data providers all state organisations that account for 98.5% of records and two of them are from the Northern Territory. The Northern Territory counts accounts for 56% of dingo records in contrast to Western Australia, Victoria, Australian Capital Territory and Queensland collectively accounting for only 5% of records. This is a heavily skewed distribution and could be explained by the legislation each state holds around dingoes. 

Attitudes towards dingoes and their conservation status vary between states/territores. For example, the Northern Territory is the only state in Australia that recognizes dingoes as a protected species, however permits can still be applied for to take or remove wildlife. In other states such as Western Australia, New South Wales, Queensland and South Australia inside the dog fence dingoes are declared a pest species. Outside the Dog Fence they are managed as unprotected native animals under the National Parks and Wildlife Act 1972 because they also have an important ecological role. These differences in legislation between states might help to explain why the highest numbers of observations are in states and territories that recognize dingoes as protected.


```{r}
Dataproviders <- ggplot() +
  geom_sf(data = aus, fill = "beige") +
  geom_point(data = points_fitered,
             mapping = aes(x = decimalLongitude,
                           y = decimalLatitude,
                           colour = dataResourceName),
             alpha = 0.25) +
  scale_color_manual(values = c(
    "Museums Victoria provider for OZCAM" = "#A838FF", 
    "Victorian Biodiversity Atlas" = "#6095FF", 
    "Australian National Wildlife Collection provider for OZCAM" = "#FE7D49", 
    "NSW BioNet Atlas" = "#FF38DE", 
    "SA Fauna (BDBSA)" = "#58D68D", 
    "Australian Museum provider for OZCAM" = "#FFC300", 
    "Fauna Atlas N.T." = "#566573")) +
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0), 
        panel.background = element_rect(fill = "white", size = 2), 
        panel.grid.major = element_line(colour = NA),
        plot.title = element_text(face = "bold", size = 16, hjust = 0.5, line = 3),
        axis.ticks = element_line(colour = "grey70", size = 0.2),
        axis.title.y = element_text(size = 12, vjust = 3),
        axis.title.x = element_text(size = 12, vjust = -3),
        legend.position = "none") +
  guides(color = guide_legend(title = "Data Providers", alpha = "none")) +
  labs(x = "Longitude",
       y = "Latitude")


Barplot <-  counts_fitered %>%
  ggplot(aes(x = reorder(str_wrap(dataResourceName, 25), - count), 
             y = count, fill = dataResourceName)) +
    geom_bar(stat="identity", width=.8) +
    scale_fill_manual(values = c(
    "Museums Victoria provider for OZCAM" = "#A838FF", 
    "Victorian Biodiversity Atlas" = "#6095FF", 
    "Australian National Wildlife Collection provider for OZCAM" = "#FE7D49", 
    "NSW BioNet Atlas" = "#FF38DE", 
    "SA Fauna (BDBSA)" = "#58D68D", 
    "Australian Museum provider for OZCAM" = "#FFC300", 
    "Fauna Atlas N.T." = "#566573")) +
    coord_flip() +
    xlab("") +
    theme_bw() + 
  theme(legend.position = "none",
        axis.text = element_text(size = 12))


# Plot bar plot and map
Barplot + Dataproviders + 
  plot_annotation(
    title = "Data Providers for Dingo Occurrences"
  ) + 
  plot_layout(widths = c(1, 2),
              heights = c(2, 1))
```

# When are observations recorded?


We can also investigate what months of the year are they been recorded and if there any patterns that can tell us about the biology of species of when they are most active or the bias in data collection when people like to go and record dingoes. With the Northern Territory and South Australia having the highest counts, it may also be good to break the times of year its recorded by the state.

```{r}
dingo <- galah_call() |>
  galah_identify("canis familiaris") |>
  galah_filter(raw_vernacularName == "Dingo") |>
  galah_apply_profile(ALA) |>
  galah_select(group = "basic", dataResourceName) |>
  atlas_occurrences()

dingo_dates <- dingo |>
  mutate(eventDate = ymd_hms(eventDate), 
         year = year(eventDate),
         month = month(eventDate, label = TRUE),
         julian = yday(eventDate))
         
month_counts <- dingo_dates|>
  group_by(month) |>
  count()

ggplot() +
  geom_bar(data = month_counts,
           mapping = aes(x = month, y = n),
           stat = "identity")

dingo_NT <- galah_call() |>
  galah_identify("canis familiaris") |>
  galah_filter(raw_vernacularName == "Dingo") |>
  galah_apply_profile(ALA) |>
  galah_select(group = "basic", dataResourceName, cl22) |>
  galah_filter(cl22 == "Northern Territory") |>
  atlas_occurrences()
dingo_NT

dingo_dates_NT <- dingo_NT |>
  mutate(eventDate = ymd_hms(eventDate), 
         year = year(eventDate),
         month = month(eventDate, label = TRUE),
         julian = yday(eventDate))

month_counts_NT <- dingo_dates_NT |>
  group_by(month) |>
  count()



NT <- ggplot() +
  geom_bar(data = month_counts_NT,
           mapping = aes(x = month, y = n),
           stat = "identity")



dingo_SA <- galah_call() |>
  galah_identify("canis familiaris") |>
  galah_filter(raw_vernacularName == "Dingo") |>
  galah_apply_profile(ALA) |>
  galah_select(group = "basic", dataResourceName, cl22) |>
  galah_filter(cl22 == "South Australia") |>
  atlas_occurrences()

dingo_dates_SA <- dingo_SA |>
  mutate(eventDate = ymd_hms(eventDate), 
         year = year(eventDate),
         month = month(eventDate, label = TRUE),
         julian = yday(eventDate))

month_counts_SA <- dingo_dates_SA |>
  group_by(month) |>
  count()

SA <- ggplot() +
  geom_bar(data = month_counts_SA,
           mapping = aes(x = month, y = n),
           stat = "identity")
```

```{r}
SA + NT
```


In South Australia and the Northern Territory, majority of counts are been recorded during winter, this is telling us that either dingoes are more active during winter or data collectors prefer surveying during cooler times of the year. Research has shown that dingoes studies in the Northern Territory, energy expenditure was significantly lower during summer with 91% of their day was spent stationary in comparison to 46% in Winter. This was measured through tri-axial accelerometry and using GPS data. 



Dingoes are the most protected within National Parks because of the permits that can be applied for to take or remove wildlife outside of these areas, therefore I decided to plot dingo observations over Nationally protected areas as well as the dingo fence within South Australia to see if there is any correlation between the legislation in these areas and the occurrence of dingoes. This is point density map which shows overlapping observations. 

Expand code here:  

```{r}
capad <- st_read(here("/Users",
                      "smi9e9",
                      "Documents",
                      "tutorials",                     
                      "CAPAD",
                      "CAPAD2020_terrestrial.shp"),
                 quiet = TRUE) |>
  ms_simplify(keep = 0.1)
```

```{r}
capad <- capad |>
  st_transform(crs = st_crs("WGS84")) |>
  st_make_valid()
```

```{r}
Dog_Fence <- st_read(here("/Users",
                      "smi9e9",
                      "Documents",
                      "tutorials",                     
                      "Dog Fence",
                      "Dog_Fence.shp"),
                 quiet = TRUE) |>
  ms_simplify(keep = 0.1)
```

```{r}
Dog_Fence <- Dog_Fence |>
  st_transform(crs = st_crs("WGS84")) |>
  st_make_valid()
```
```{r}
galah_config(email = "SMI9E9@csiro.au", verbose = FALSE)
```

```{r}
dingoocc <- galah_call() |>
  galah_identify("canis familiaris") |>
  galah_filter(raw_vernacularName == "Dingo") |>
  galah_apply_profile(ALA) |>
  galah_group_by(dataResourceName, cl22) |>
  atlas_occurrences()
```
 
```{r}
dingo_palette <- list(
                      "Protected" = "#00A600FF",
                      "Areas" = "#00A600FF",
                      "Dingo" = "#67136B",
                      "observations" = "#FC6B00")
```

```{r}
dingo_records <- galah_call() |>
  galah_identify("canis familiaris") |>
  galah_filter(raw_vernacularName == "Dingo")|>
  atlas_occurrences()
```

```{r}
# Tidy
dingo_records <- dingo_records |>
  drop_na() |>                      # drop NAs
  filter(decimalLongitude < 155,    # filter records to only mainland/tasmania
         decimalLongitude > 110, 
         decimalLatitude > -45,
         decimalLatitude < -10)
```

```{r}
NT_plot <- ggplot() +
  geom_sf(data = ozmap_states, 
          fill = "beige", colour = "black", linewidth = 0.3) +
  geom_sf(data = capad, fill = "#00A600FF", colour = "black", linewidth = 0.25, alpha = 0.3, linetype = "dashed") +
  geom_pointdensity(data = dingo_records, size = 0.7, alpha = 0.35,
                    mapping = aes(x = decimalLongitude,
                                  y = decimalLatitude)) +
  viridis::scale_color_viridis(option = "B", 
                              begin = 0.1, 
                             guide_colorbar( 
                                            title = ("Number of \nOverlapping observations"), 
                                            alpha = "none")) +
  geom_sf(data = Dog_Fence, color = "blue", linewidth = 1) +
  coord_sf(xlim = c(128, 139), ylim = c(-26, -10)) + # filter to mainland australia+
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0), 
        panel.background = element_rect(fill = "white", size = 2), 
        panel.grid.major = element_line(colour = NA),
        legend.position = "bottom",
        axis.ticks = element_line(colour = "grey70", size = 0.2),
        axis.title.y = element_text(size = 12),
        axis.title.x = element_text(size = 12)) 
```

```{r}
NT_plot
```

```{r}
NT_plot <- NT_plot +
 ggnewscale::new_scale_color() +
 scale_colour_manual(values = dingo_palette) +
  labs(title = glue("Dingo observations in <br> Australian <span style='color:{dingo_palette$Protected}'>**Protected**</span> <span style='color:{dingo_palette$Areas}'>**Areas**</span> <br> in the Northern Territory <br>")) +
 theme(plot.title = element_markdown(face = "bold", size = 16, hjust = 0.5)) +
annotate(geom = "line", colour = "blue", linewidth = 1) +        
 annotate(geom = "text", label = "Dingo Fence", hjust = "left") +#couldnt get the annotation to work 
  labs(y = "Latitude", x = "Longitude") 
```

```{r}
NT_plot
```
```{r}
SA_plot <- ggplot() +
  geom_sf(data = ozmap_states, 
          fill = "beige", colour = "black", linewidth = 0.3) +
  geom_sf(data = capad, fill = "#00A600FF", colour = "black", linewidth = 0.25, alpha = 0.3, linetype = "dashed") +
  geom_pointdensity(data = dingo_records, size = 0.7, alpha = 0.35,
                    mapping = aes(x = decimalLongitude,
                                  y = decimalLatitude)) +
    viridis::scale_color_viridis(option = "B", 
                               begin = 0.1,
                               guide_colorbar(
                                              title = ("Number of \nOverlapping observations"), 
                                              alpha = "none")) +
  geom_sf(data = Dog_Fence, color = "blue", linewidth = 1) +
  coord_sf(xlim = c(128, 141), ylim = c(-26, -40)) + # filter to mainland australia+
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0), 
        panel.background = element_rect(fill = "white", size = 2), 
        panel.grid.major = element_line(colour = NA),
        legend.position = "bottom",
        axis.ticks = element_line(colour = "grey70", size = 0.2),
        axis.title.y = element_text(size = 12),
        axis.title.x = element_text(size = 12)) 
```

```{r}
SA_plot
```

```{r}
SA_plot2 <- SA_plot +
  ggnewscale::new_scale_color() +
  scale_colour_manual(values = dingo_palette) +
  labs(title = glue("Dingo observations in <br> Australian <span style='color:{dingo_palette$Protected}'>**Protected**</span> <span style='color:{dingo_palette$Areas}'>**Areas**</span> <br> in South Australia <br>")) +
  theme(plot.title = element_markdown(face = "bold", size = 16, hjust = 0.5)) +
  annotate(geom = "line", colour = "blue", linewidth = 1) +        
  annotate(geom = "text", label = "Dingo Fence", hjust = "left") + #couldnt get the annotation to work 
  labs(y = "Latitude", x = "Longitude")
```
 
```{r}
SA_plot2
```

In this map, it is visually quite obvious especially in the Northern Territory that majority of the counts fall within the Nationally protected areas. In South Australia, it is also clear dingoes are only been recorded on one side of the fence where they remain an unprotected native species. Therefore, we can see there is correlation between legislation effecting the data collection and the occurence of a species. 

In conclusion, dingoes are ecologically important playing an important role as apex predators, regulating ecological cascades. The impact they have on the sheep industry is recognised, however because of their ecological importance we need to find a balance in keeping this species protected in areas for maintaining healthy ecosystems, where native small mammals can persist. The data collection of a species becomes more valuable when there public invested interest and we see this a species we want to protect. Just relying on state organisation to record dingo observations limits our knowledge on the distribution of the dingo in Australia and where they actually are now. 





