---
title: "Christmas Beetles"
author: "Dax Kellie"
date: "4 December, 2023"
title-block-banner: "#B8573E"
format: 
  html: 
    fontfamily: Roboto
    mainfont: Roboto
    sansfont: Roboto
    monofont: Fira Code
    linkcolor: "#E06E53"
    highlight-style: arrow
editor_options: 
  chunk_output_type: console
---

This is an investigation of Christmas Beetle records in the ALA.

```{r setup}
#| include: false
#| message: false
#| warning: false

# packages
library(galah)
library(tidyverse)
library(here)
library(sf)
library(ozmaps)
library(showtext)
library(ggtext)

galah_config(email = "dax.kellie@csiro.au")

font_add_google("Roboto Condensed", "Roboto")
showtext_auto(enable = TRUE)

# create a list of all years
all_years <- tibble(year = c(1900:2023))

# Map of australia
aus <- ozmap_states |>
  st_transform(crs = st_crs(4326))
```

# Background

## Beetles

Counts of all beetles in 2023:

```{r}
beetle_taxon <- search_taxa("coleoptera")

beetles <- galah_call() |>
  identify(beetle_taxon$taxon_concept_id) |>
  filter(year > 1900) |>
  group_by(year) |>
  atlas_counts() |>
  mutate(year = as.numeric(year)) |>
  right_join(all_years, join_by(year == year)) |>
  replace_na(list(count = 0))

beetles |>
  ggplot(aes(x = year, y = count)) + 
  geom_bar(stat = "identity", fill = "#A191B2") + 
  labs(title = "<i>Coleoptera</i>") +
  pilot::theme_pilot(axes = "l",
                     grid = "h") + 
  theme(plot.title = element_markdown(family = "Roboto", halign = 0.5))
```

## Scarab beetles

Counts of all scarab beetles (family *Scarabaeidae*)

```{r}
scarabs <- galah_call() |>
  identify("Scarabaeidae") |>
  filter(year > 1900) |>
  group_by(year) |>
  atlas_counts() |>
  mutate(year = as.numeric(year)) |>
  right_join(all_years, join_by(year == year)) |>
  replace_na(list(count = 0)) 

scarabs |>
  ggplot(aes(x = year, y = count)) + 
  geom_bar(stat = "identity", fill = "#A191B2") + 
  labs(title = "*Scarabaeidae*") +
  pilot::theme_pilot(axes = "l",
                     grid = "h") + 
  theme(plot.title = element_markdown(family = "Roboto", halign = 0.5))
```

# Christmas beetles

## Counts

Counts of Christmas beetles (genus *Anoplognathus*)

Total:

```{r}
c_beetle_taxon <- search_taxa("Anoplognathus Leach, 1815")

galah_call() |>
  identify(c_beetle_taxon$taxon_concept_id) |>
  filter(year > 1900) |>
  atlas_counts()
```

Total before 2022:

```{r}
# total before 2022
galah_call() |>
  identify(c_beetle_taxon$taxon_concept_id) |>
  filter(year < 2022) |>
  atlas_counts()
```


```{r}
# total before 2022
galah_call() |>
  identify(c_beetle_taxon$taxon_concept_id) |>
  atlas_counts()
```

**% Growth since Christmas Beetle count initiative in 2022**

```{r}
# growth in observations due to christmas beetle count
(16655/12719) * 100
```


```{r}
c_beetles <- galah_call() |>
  identify(c_beetle_taxon$taxon_concept_id) |>
  filter(year > 1899) |>
  group_by(year) |>
  atlas_counts() |>
  mutate(year = as.numeric(year)) |>
  right_join(all_years, join_by(year == year)) |>
  replace_na(list(count = 0))

c_beetles |>  
  ggplot(aes(x = year, y = count)) + 
  geom_bar(stat = "identity", fill = "#A191B2") + 
  labs(title = "Christmas Beetles") +
  pilot::theme_pilot(axes = "l",
                     grid = "h") + 
  theme(plot.title = element_markdown(family = "Roboto", halign = 0.5))
```

Christmas beetle counts scaled by all beetle counts

```{r}
# Scale christmas beetle counts
c_beetles_scaled <- c_beetles |>
  left_join(scarabs, join_by(year == year), suffix = c("", "_scarabs")) |>
  left_join(beetles, join_by(year == year), suffix = c("", "_beetles")) |>
  mutate(
    prop_of_scarabs = count / count_scarabs,
    prop_of_beetles = count / count_beetles
  ) %>%
  mutate_if(is.numeric, ~ replace(., is.infinite(.), 0))

c_beetles_scaled |>  
  ggplot() + 
  # geom_bar(aes(x = year, y = count), stat = "identity", fill = "#A191B2") + 
  geom_smooth(aes(x = year, y = prop_of_beetles), method="loess",
            color = pilot::pilot_color("orange")) +
  labs(title = "Proportion of christmas beetle observations to all beetle observations") +
  pilot::theme_pilot(axes = "l",
                     grid = "h") + 
  theme(plot.title = element_markdown(family = "Roboto", halign = 0.5))
```

Ok so it turns out christmas beetles are fairly well recorded in the last 2 years, which means scaling counts likely won't really control for growth in overall counts (they are driving this growth in beetles)

```{r}
galah_call() |>
  identify("coleoptera") |>
  filter(year > 2020) |>
  group_by(year, genus) |>
  atlas_counts(limit = 5)
```


Scaled by total insects, observations still appear to be going up (though the actual number is tiny)

```{r}
insects <- galah_call() |>
  identify("insecta") |>
  filter(year > 1899) |>
  group_by(year) |>
  atlas_counts() |>
  mutate(year = as.numeric(year)) |>
  right_join(all_years, join_by(year == year)) |>
  replace_na(list(count = 0))

c_beetles_scaled <- c_beetles_scaled |>
  left_join(insects, join_by(year == year), suffix = c("", "_insects")) |>
  mutate(
    prop_of_insects = count / count_insects,
  ) %>%
  mutate_if(is.numeric, ~ replace(., is.infinite(.), 0))

c_beetles_scaled |>
  filter(year > 2000) |>
  ggplot() + 
  # geom_bar(aes(x = year, y = count), stat = "identity", fill = "#A191B2") + 
  geom_smooth(aes(x = year, y = prop_of_insects), method="loess",
            color = pilot::pilot_color("orange")) +
  labs(title = "Proportion of christmas beetle observations to all insect observations") +
  pilot::theme_pilot(axes = "l",
                     grid = "h") + 
  theme(plot.title = element_markdown(family = "Roboto", halign = 0.5))
```

## Observations

```{r}
# counts by data resource
galah_call() |>
  identify(c_beetle_taxon$taxon_concept_id) |>
  group_by(dataResourceName) |>
  galah_apply_profile(ALA) |>
  atlas_counts()

# observations
c_beetle_obs <- galah_call() |>
  identify(c_beetle_taxon$taxon_concept_id) |>
  galah_apply_profile(ALA) |>
  select(group = "basic", year, dataResourceName) |>
  atlas_occurrences()
```

```{r}
# argentinian scarabs
ggplot() + 
  geom_sf(data = aus,
          fill = "white",
          colour = "grey60") + 
  geom_point(data = c_beetle_obs |>
               filter(dataResourceName == "iNaturalist Australia",
                      year > 2000),
             aes(x = decimalLongitude,
                 y = decimalLatitude),
             colour = pilot::pilot_color("navy"),
             alpha = 0.6
             ) +
  geom_point(data = c_beetle_obs |>
               filter(dataResourceName != "iNaturalist Australia",
                      year > 2000),
             aes(x = decimalLongitude,
                 y = decimalLatitude),
             colour = pilot::pilot_color("orange"),
             alpha = 0.6
             ) +
  coord_sf(xlim = c(110, 155), 
           ylim = c(-45, -10)) +
  labs(title = "Christmas beetle observations since 2000",
       subtitle = glue::glue("<span style = 'color: #204466;'>**iNaturalist**</span> vs 
                             <span style = 'color: #f28100;'>all other observations</span>")) +
  theme_void() + 
  theme(
    plot.title = ggtext::element_markdown(family = "Roboto", hjust = 0.5),
    plot.subtitle = ggtext::element_markdown(family = "Roboto", hjust = 0.5)
  )
```



# Plants

What about plants that christmas beetles rely on for food? Have those observations decreased?


#### Eucalypts

```{r}
eucalypt_taxon <- search_taxa("eucalyptus")

eucalypts <- galah_call() |>
  identify(eucalypt_taxon$taxon_concept_id) |>
  filter(year > 1899) |>
  group_by(year, species) |>
  atlas_counts() |>
  mutate(year = as.numeric(year)) |>
  right_join(all_years, join_by(year == year)) |>
  replace_na(list(count = 0))

eucalypt_total <- galah_call() |>
  identify(eucalypt_taxon$taxon_concept_id) |>
  filter(year > 1899) |>
  group_by(year) |>
  atlas_counts() |>
  mutate(year = as.numeric(year)) |>
  right_join(all_years, join_by(year == year)) |>
  replace_na(list(count = 0))

euc_scaled <- eucalypts |>
  left_join(eucalypt_total, join_by(year == year), suffix = c("", "_total")) |>
  mutate(
    prop_of_euc = count / count_total,
  ) %>%
  mutate_if(is.numeric, ~ replace(., is.infinite(.), 0))

euc_scaled |>
  filter(str_detect(species, "Eucalyptus melliodora"))

ggplot() + 
  geom_line(data = euc_scaled,
            aes(x = year, y = prop_of_euc, group = species),
            colour = "grey80",
            stat = "identity") +
  geom_line(data = euc_scaled |> filter(species == "Eucalyptus camaldulensis"),
            aes(x = year, y = prop_of_euc),
            stat = "identity",
            colour = pilot::pilot_color("purple"),
            linewidth = 1.3) +
  geom_line(data = euc_scaled |> filter(species == "Eucalyptus melliodora"),
            aes(x = year, y = prop_of_euc),
            stat = "identity",
            colour = pilot::pilot_color("orange"),
            linewidth = 1.3) +
  geom_line(data = euc_scaled |> filter(species == "Eucalyptus sideroxylon"),
            aes(x = year, y = prop_of_euc),
            stat = "identity",
            colour = pilot::pilot_color("green"),
            linewidth = 1.3) +
  geom_line(data = euc_scaled |> filter(species == "Eucalyptus conica"),
            aes(x = year, y = prop_of_euc),
            stat = "identity",
            colour = pilot::pilot_color("blue"),
            linewidth = 1.3) +

  # geom_smooth(aes(x = year, y = prop_of_euc, fill = species), method="loess") +
  labs(title = "Proportion of eucalypt observations") +
  # pilot::scale_color_pilot() +
  pilot::theme_pilot(axes = "l",
                     grid = "h") + 
  theme(
    legend.position = "none",
    plot.title = element_markdown(family = "Roboto", halign = 0.5)
    )
```

### Native grasses

```{r}
search_taxa(c("Chloridoideae", "Poaceae", "Pooideae"))

galah_call() |>
  identify("Chloridoideae", "Poaceae", "Pooideae") |>
  group_by(year) |>
  atlas_counts() |>
  
  ggplot() +
  geom_bar(aes(x = year, y = count), stat = "identity", fill = pilot::pilot_color("green")) +
  labs(title = "Grasses counts") +
  pilot::theme_pilot(axes = "l",
                     grid = "h")
```


# Christmas beetles misidentification

Top 3 misIDs are:

  1.  Argentinian scarab (Cyclocephala signaticollis)
  2.  June beetle (Melolonthinae, assorted species)
  3.  Golden stag beetle (Lamprima aurata)

```{r}
mis_id_taxon <- search_taxa(c("Cyclocephala signaticollis", 
                              "Melolonthinae", 
                              "Lamprima aurata"
                              ))
mis_id_taxon
```

## Counts

There is a growth in these 3 commonly misID'd beetles, particularly since the Christmas Beetle count was started

```{r}
arg_scarab <- galah_call() |>
  identify(mis_id_taxon$taxon_concept_id[1]) |>
  filter(year > 1999) |>
  group_by(year) |>
  atlas_counts() |>
  mutate(year = as.numeric(year),
         name = rep("Argentinian scarab beetle")) |>
  right_join(all_years, join_by(year == year)) |>
  replace_na(list(count = 0))

june_beetle <- galah_call() |>
  identify(mis_id_taxon$taxon_concept_id[2]) |>
  filter(year > 1999) |>
  group_by(year) |>
  atlas_counts() |>
  mutate(year = as.numeric(year),
         name = rep("June beetle (assorted species)")) |>
  right_join(all_years, join_by(year == year)) |>
  replace_na(list(count = 0))

golden_stag <- galah_call() |>
  identify(mis_id_taxon$taxon_concept_id[3]) |>
  filter(year > 1999) |>
  group_by(year) |>
  atlas_counts() |>
  mutate(year = as.numeric(year), 
         name = rep("Golden stag beetle")) |>
  right_join(all_years, join_by(year == year)) |>
  replace_na(list(count = 0))

mis_id_joined <- arg_scarab |>
  full_join(june_beetle) |>
  full_join(golden_stag) |>
  drop_na(name)

ggplot(data = mis_id_joined) +
  geom_bar(aes(x = year, y = count, fill = name), stat = "identity") + 
  pilot::scale_fill_pilot() +
  labs(title = "Observations of common misID'd species believed to be Christmas beetles") +
  pilot::theme_pilot(axes = "l",
                     grid = "h") + 
  theme(
    legend.position = "none",
    plot.title = element_markdown(family = "Roboto", halign = 0.5)
    )  +
  facet_grid( ~ name)

```

## Observations

Observations of beetles has grown, and so too has the distribution - especially of the 2 scarab beetles. This growth has implications for our understanding of these species, and it particularly relevant for Argentinian scarab beetles, which are invasive.

```{r}
# argentinian scarabs

# counts by data resource
galah_call() |>
  identify(mis_id_taxon$taxon_concept_id[1]) |>
  group_by(dataResourceName) |>
  galah_apply_profile(ALA) |>
  atlas_counts()

# observations
arg_scarab_obs <- galah_call() |>
  identify(mis_id_taxon$taxon_concept_id[1]) |>
  select(group = "basic", year, dataResourceName) |>
  galah_apply_profile(ALA) |>
  atlas_occurrences()

search_all(fields, "basis") |> show_values()
search_all(fields, "dataResourceName") |> search_values("museum")
```


```{r}
# june beetles

# counts by data resource
galah_call() |>
  identify(mis_id_taxon$taxon_concept_id[2]) |>
  group_by(dataResourceName) |>
  galah_apply_profile(ALA) |>
  atlas_counts()

# observations
june_beetle_obs <- galah_call() |>
  identify(mis_id_taxon$taxon_concept_id[2]) |>
  galah_apply_profile(ALA) |>
  select(group = "basic", year, dataResourceName) |>
  atlas_occurrences()
```


```{r}
# golden stags

# counts by data resource
galah_call() |>
  identify(mis_id_taxon$taxon_concept_id[3]) |>
  group_by(dataResourceName) |>
  galah_apply_profile(ALA) |>
  atlas_counts()

# observations
golden_stag_obs <- galah_call() |>
  identify(mis_id_taxon$taxon_concept_id[3]) |>
  select(group = "basic", year, dataResourceName) |>
  galah_apply_profile(ALA) |>
  atlas_occurrences()
```


```{r}

# argentinian scarabs
ggplot() + 
  geom_sf(data = aus,
          fill = "white",
          colour = "grey60") + 
  geom_point(data = arg_scarab_obs |>
               filter(dataResourceName == "iNaturalist Australia",
                      year > 2000),
             aes(x = decimalLongitude,
                 y = decimalLatitude),
             colour = pilot::pilot_color("navy"),
             alpha = 0.6
             ) +
  geom_point(data = arg_scarab_obs |>
               filter(dataResourceName != "iNaturalist Australia",
                      year > 2000),
             aes(x = decimalLongitude,
                 y = decimalLatitude),
             colour = pilot::pilot_color("orange"),
             alpha = 0.6
             ) +
  labs(title = "Argentinian scrab beetle observations since 2000",
       subtitle = glue::glue("<span style = 'color: #204466;'>**iNaturalist**</span> vs 
                             <span style = 'color: #f28100;'>all other observations</span>")) +
  theme_void() + 
  theme(
    plot.title = ggtext::element_markdown(family = "Roboto", hjust = 0.5),
    plot.subtitle = ggtext::element_markdown(family = "Roboto", hjust = 0.5)
  )


# june beetles
ggplot() + 
  geom_sf(data = aus,
          fill = "white",
          colour = "grey60") + 
  geom_point(data = june_beetle_obs |>
               filter(dataResourceName == "iNaturalist Australia",
                      year > 2000),
             aes(x = decimalLongitude,
                 y = decimalLatitude),
             colour = pilot::pilot_color("navy"),
             alpha = 0.6
             ) +
  geom_point(data = june_beetle_obs |>
               filter(dataResourceName != "iNaturalist Australia",
                      year > 2000),
             aes(x = decimalLongitude,
                 y = decimalLatitude),
             colour = pilot::pilot_color("orange"),
             alpha = 0.6
             ) +
  coord_sf(xlim = c(110, 155), 
           ylim = c(-45, -10)) +
  labs(title = "June beetle observations since 2000",
       subtitle = glue::glue("<span style = 'color: #204466;'>**iNaturalist**</span> vs 
                             <span style = 'color: #f28100;'>all other observations</span>")) +
  theme_void() + 
  theme(
    plot.title = ggtext::element_markdown(family = "Roboto", hjust = 0.5),
    plot.subtitle = ggtext::element_markdown(family = "Roboto", hjust = 0.5)
  )  

# golden stag beetles
ggplot() + 
  geom_sf(data = aus,
          fill = "white",
          colour = "grey60") + 
  geom_point(data = golden_stag_obs |>
               filter(dataResourceName == "iNaturalist Australia",
                      year > 2000),
             aes(x = decimalLongitude,
                 y = decimalLatitude),
             colour = pilot::pilot_color("navy"),
             alpha = 0.6
             ) +
  geom_point(data = golden_stag_obs |>
               filter(dataResourceName != "iNaturalist Australia",
                      year > 2000),
             aes(x = decimalLongitude,
                 y = decimalLatitude),
             colour = pilot::pilot_color("orange"),
             alpha = 0.6
             ) +
  coord_sf(xlim = c(110, 155), 
           ylim = c(-45, -10)) +
  labs(title = "Golden stag beetle observations since 2000",
       subtitle = glue::glue("<span style = 'color: #204466;'>**iNaturalist**</span> vs 
                             <span style = 'color: #f28100;'>all other observations</span>")) +
  theme_void() + 
  theme(
    plot.title = ggtext::element_markdown(family = "Roboto", hjust = 0.5),
    plot.subtitle = ggtext::element_markdown(family = "Roboto", hjust = 0.5)
  )  

```


## Invasive species

```{r}
library(readxl)

taxa_mis_id <- readxl::read_excel(here("projects", 
                                       "christmas-beetles", 
                                       "data",
                                       "xmas-beetle-confusion.xlsx"
                                       ),
                                  sheet = 1, 
                                  range = cell_cols("A:C")) |>
  janitor::clean_names()

taxa_list <- taxa_mis_id |>
  pivot_longer(cols = 1:3,
               names_to = "mis_id_type") |>
  rename(name = value) |> 
  drop_na(name)

matched_taxa <- search_taxa(c(taxa_list$name)) |>
  dplyr::filter(!is.na(taxon_concept_id))


# counts plot

get_counts <- function(taxon_id) {
  
  taxon_name <- search_taxa(taxon_id) |> 
    dplyr::select(scientific_name) |>
    pull(scientific_name)
  
  year_range <- tibble(year = c(2000:2023))
  
  query <- galah_call() |>
  identify(taxon_id) |>
  filter(year > 1999) |>
  group_by(year) |>
  atlas_counts() |>
  mutate(year = as.numeric(year)) |>
  right_join(year_range, join_by(year == year)) |>
  replace_na(list(count = 0)) |>
  mutate(name = rep(taxon_name))
  
}

year_counts <- matched_taxa |>
  select(taxon_concept_id) |>
  pull(taxon_concept_id) |>
  map(get_counts)

year_counts <- year_counts |> bind_rows()

ggplot(data = year_counts |>
         filter(!name %in% c("CARABIDAE", "Onthophagus"))) +
  geom_bar(aes(x = year, y = count, fill = name), stat = "identity") + 
  pilot::scale_fill_pilot() +
  labs(title = "Observations of common misID'd species believed to be Christmas beetles") +
  pilot::theme_pilot(axes = "l",
                     grid = "h") + 
  theme(
    legend.position = "none",
    plot.title = element_markdown(family = "Roboto", halign = 0.5)
    )  +
  facet_wrap( ~ name)
```

```{r}
# check whether any are on epbc list
galah_call() |>
  identify(matched_taxa$taxon_concept_id) |>
  atlas_species()

search_all(lists, "threatened") |>
  dplyr::filter(isThreatened == TRUE)

epbc_list <- search_all(lists, "dr656") |>
  show_values()

matched_taxa |>
  select(scientific_name) |> print(n = 62)
```

