---
title: Magpies
description: |
  Visualisation of mapgpie occurrence records in the ALA
author:
  - name: Matilda Stevenson
    url: "people/Stevenson_Matilda/index.html"
date: '2021-04-14'
categories: 
  - Magpies
  - Data Visualisation
output: 
  distill::distill_article:
    highlight: tango
    toc: true
    toc_float: true
    toc_depth: 3
---

The Atlas of Living Australia (ALA) holds records of magpie sightings from a number data providers like iNaturalist, eBird and BirdLife AUstralia. We wanted to make a visualisation of [Australian Bird of the Year 2018 winner](https://www.abc.net.au/news/2017-12-11/bird-of-the-year-magpie-defeats-teambinchicken/9245242), Magpies, using records held in the ALA.

<aside>
```{r author image, echo = FALSE, eval = FALSE}
library(here)
image_path <- here("comms", "images", "matilda.jpg")
htmltools::withTags(
# Save your own picture as "picture.jpg" to local directory
      img(src = knitr::image_uri(image_path),
          style = 'height:120px;clip-path: circle();'))
```
</aside>


See the [original R script](https://github.com/AtlasOfLivingAustralia/science/blob/main/comms/2021-04_magpies/magpie_occ_vis.R)



***

# Download records

First we load the required packages

```{r, message = FALSE, warning = FALSE}
# packages
library(ggplot2)
library(tidyr)
library(ozmaps)
library(sf)
# library(hexbin) # You should be prompted to download when creating map
```

We will use the [galah](https://atlasoflivingaustralia.github.io/galah/index.html) package to download records.
(If running this code yourself, you also need to set an email associated with your data download request, substituting your email with `myemail@email.com`)

```{r, eval = FALSE}
library(galah)
ala_config(email = "myemail@email.com")
```

Now we can download magpie occurrence records by using `ala_occurrences()`

```{r, results = 'hide'}
# Download magpie occurrence records using `galah`
magpie_occ <- ala_occurrences(
  taxa = select_taxa("Cracticus tibicen"),
  filters = select_filters(profile = "ALA")
)
```

Filter records not on the mainland or Tasmania

```{r}
filtered_occ <- magpie_occ %>% dplyr::filter(decimalLongitude < 155,
                                      decimalLongitude > 110,
                                      decimalLatitude > -45,
                                      decimalLatitude < -10)
```

# Create map

```{r, eval = FALSE, echo = FALSE}
# Note: This doesn't work because hexbin does not install correctly as of 21 July, 2021

# Hexagon map using geom_hex
ggplot(filtered_occ) +
  geom_hex(mapping = aes(x = decimalLongitude, y = decimalLatitude), bins = 55) +
  ylim(-45, -10) +
  xlim(110, 155) +
  scale_fill_gradientn(colours = c("#EEECEA", "#E06E53")) +
  theme_bw() +
  theme(
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    legend.position = "none")
```

We can build a hexagon map by manually creating hexagons, linking them to a grid of Australia and adding our data of magpie counts to colour the hexagons.

First we convert the aus map from ozmaps to the same coordinate system as ALA data

```{r}
# Convert map of aus to same coordinate system as ALA points
aus <- st_transform(ozmaps::ozmap_country, 4326)
```

Then transform points from our data frame of magpie occurrences to the correct coordinates

```{r}
# transform points
points <- lapply(1:nrow(filtered_occ), function(x) {
  st_point(c(filtered_occ[x,'decimalLongitude'],
             filtered_occ[x,'decimalLatitude']))
})
```

Now build the hexagonal grid

```{r}
# Build hexagonal grid
grid_all <- st_make_grid(aus, cellsize = 1, what = "polygons", square = FALSE)[aus]
gridSF <- st_as_sf(grid_all)

# Find which polygon each point is in
intersect <- st_intersects(st_as_sfc(points) %>% st_set_crs(4326), gridSF)

# count the number of points in each hexagon
freqs <- as.data.frame(table(unlist(intersect)), stringsAsFactors = FALSE)

names(freqs) <- c("hex_id", "count")
```

Add our magpie counts data to the hexagons and fill hexagons without data

```{r}
# Add counts to hexagons
freqs$hex_id <- as.integer(freqs$hex_id)

# fill in the values for hexagons with no records
filled_freqs <- freqs %>%
  complete(hex_id = full_seq(c(1,925), 1), fill = list(count = 0))
gridSF$n <- filled_freqs$count
```

Finally, build the plot

```{r, message = FALSE, message = FALSE, layout = "l-page",preview = TRUE}
# Plot the hexagons in ALA colours, with log-transformed counts
ggplot() +
  geom_sf(data = gridSF, aes(fill = n), size = .01) +
  scale_fill_gradientn(colours = c("#EEECEA", "#E06E53"), na.value = "white", trans = "log2") +
  ggtitle("Magpies in Australia", "Based on data of magpie occurrences in the ALA") +
  ylim(-45, -10) +
  xlim(110, 155) +
  theme_bw() +
  theme(
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    legend.position = "none") 
```


