---
title: "19/23 DOI"
format: html
editor: visual
---

# Step 1 - Data cleaning:

1.  Load required packages

```{r}
#| message: false
library(here)
library(tidyverse)
library(janitor)
library(lubridate)
```

2.  Load in exported .csv from Zotero

First you will need to export the latest encompassing .csv from Zotero and save it on your local system. You can do this by selecting **File \> Export library** in the desktop application.

```{r}
#| message: false
#| warning: false
data <- read_csv(here("projects", "literature-tracking", "data", "7aug23.csv"))
```

3.  Clean column names - it will make it easier to process and interpret the data.

```{r}
dataclean <- janitor ::clean_names(data) 
```

4.  We also want to make sure the date is the same across every entry so that it processes correctly when we filter it. `as_date` standardises all date data to `yyyy-mm-dd`. We also add a new column that extracts just the month and year for the month by month analysis below.

```{r}
dataclean2 <- dataclean |>
  mutate(
    date_added_clean = as_date(date_added),
    date_added_month = month(date_added_clean),
    date_added_year = year(date_added_clean)) 
dataclean2
```

#  Step 2 - Filter 

```{r}
publicationbyyear <- dataclean2 |>
  filter(item_type == "journalArticle") |>
  filter(publication_year == c("2019", "2020", "2021", "2022")) |>
  select(date, publication_year, author, title, publication_title, doi, url) |>
  arrange(date)
  
```

#  Step 3 - Save .csv 

```{r}
publicationbyyear |>
write_csv(here("1922DOIs.csv"))
```
