---
title: "Publication tracker reporting"
toc: true
---

**Advisory Board: Quarterly Review**

Every four months the ALA is required to provide the Advisory Board an overview of major activities and outcomes. This includes a short section of \~250 words that summarises the research impact over this time.\
\
**The total amount of publications with any mention, citation or use of the ALA are provided in the quarterly report as well as the amount both citing and using the ALA.**

The quarterly review is generally updated [here](https://csiroau.sharepoint.com/:w:/r/sites/AtlasofLivingAustralia-LeadershipTeam/_layouts/15/Doc.aspx?sourcedoc=%7BD5030E16-B726-452B-B044-2E68CCE89D86%7D&file=03%20-%20ALA_Advisory_Board_18_Item_2_Directors_Update.docx&action=default&mobileredirect=true), and time of update is at request:

> This code was last updated on 2023-07-17.

> The last report looked at publications added between 2023-03-09 (9th of March 2023) and 2023-07-04 (4th of July 2023) with a publication date within 2023. Therefore, the next quarterly review will start from the **5th of July 2023**.

# Step 1: Data cleaning

1.  Load required packages

```{r, eval = FALSE}
library("here")
library("tidyverse")
library("janitor")
library("lubridate")
```

2.  Load in exported .csv from Zotero

First you will need to export the latest encompassing .csv from Zotero and save it on your local system. You can do this by selecting **File \> Export library** in the desktop application.

```{r, eval = FALSE}
data <- read_csv(here("projects", "literature-tracking", "data", "30thjune.csv"))
```

3.  Clean column names - it will make it easier to process and interpret the data.

```{r, eval = FALSE}
dataclean <- janitor ::clean_names(data) 
dataclean
```

4.  We also want to make sure the date is the same across every entry so that it processes correctly when we filter it. `as_date` standardises all date data to `yyyy-mm-dd`. We also add a new column that extracts just the month and year for the month by month analysis below.\

```{r, eval = FALSE}
dataclean2 <- dataclean |>
  mutate(
    date_added_clean = as_date(date_added),
    date_added_month = month(date_added_clean),
    date_added_year = year(date_added_clean)) |>
   select(title, date, place, publication_year, item_type, url, manual_tags, date_added, date_added_clean, date_added_month, date_added_year)
dataclean2
```

# Step 2: Summaries

## Month by month

> This isn't usually required in the quarterly review, but it may be requested at times.

1.  Here, we filter all publications added after a set date (dd/mm/yyyy) to a select date (dd/mm/yyyy). We also restrict to only those that are published in 2022+23 (sometimes old papers come in later in the tracker, we aren't really interested in these as they aren't recent).\

    ```{r, eval = FALSE}
    publicationbymonth <- dataclean2 %>%
      filter(date_added_clean >= ymd("2023-03-09")) |>
      filter(date_added_clean <= ymd("2023-07-04")) |>
      filter(publication_year == c("2023")) |>
      group_by(date_added_month, date_added_year) %>%
      count() 
    publicationbymonth
    ```

2.  Sum all publications by month

```{r, eval = FALSE}
total_sum <- sum(publicationbymonth$n) |>
print()
```

## Tags

1.  Now let's analyse this by tags (1- ALA used and 2 - ALA cited are the most important here). We'll use the above \`dataclean2\` dataset.

```{r, eval = FALSE}
dataclean3 <- dataclean2 |> 
  filter(date_added_clean >= ymd("2023-03-09")) |>
  filter(date_added_clean <= ymd("2023-07-04")) |>
  filter(publication_year == c("2023")) 
```

\
2. We need to clean the tags as they have white space and numbers that will interfere with our filtering.\
First we remove the numbers, create a new column named tags and remove the prefixes.

```{r, eval = FALSE}
remove <- c("1 - ", "2 - ", "3 - ", "4 - ", "5 - ", "6 - ") # list numbers you want removed

dataclean4 <- dataclean3 %>% # remove numbers
  mutate( # make new column called "tags"
    tags = str_remove_all(dataclean3$manual_tags, # remove prefix
                          paste(remove, collapse = "|")) 
  )
```

3.  Second, we place these tags into separate rows for the same paper. Tag names are split where semicolons are, and the `unnest` function makes them occupy multiple rows.

```{r, eval = FALSE}
## Filtering Tags

# separate manual tags into separate rows (multiple entries for one paper if inc. many tags)
ALA_tags <- dataclean4 %>%
  mutate(tags = strsplit(as.character(tags), ";")) %>% # split names where there is a semicolon
  unnest(tags) # make them occupy multiple rows

# trim whitespace from tags
ALA_tags <- ALA_tags %>% # remove numbers
  mutate(
    tags = str_trim(tags) # trim whitespace around string 
  )
```

## Publications using ALA

```{r, eval = FALSE}
ALA_used <- ALA_tags %>%
  filter(tags == "ALA used") 
ALA_used
```

## Publications

a)  All publications citing the ALA:

```{r, eval = FALSE}
ALA_cited <- ALA_tags |>
  filter(tags == "ALA cited")
ALA_cited
```

b)  All publications using the ALA:

```{r, eval = FALSE}
usedjournals <- ALA_tags |>
  filter(tags == "ALA used")
usedjournals
```

## Journal articles citing/using ALA

Publications that are journal articles:

c)  Citing the ALA

```{r, eval = FALSE}
citedjournals <- ALA_tags %>%
  filter(tags == "ALA cited") |>
  filter(item_type == "journalArticle") 
usedjournals
```

d)  Using the ALA

```{r, eval = FALSE}
citedjournals <- ALA_tags %>%
  filter(tags == "ALA used") |>
  filter(item_type == "journalArticle") 
usedjournals
```

# Open select publications

If you want to open a select handful of publications in your browser quickly, you can do so by using this code:

```{r, eval = FALSE}
usedjournals |>
  mutate(url = map(url, browseURL)) #this code will take all of the url's in the url column and open them in your default browser (NA's excluded, so be sure to check those through Zotero)
```

# Save clean .csv

Save the cleaned .csv with this code:

```{r, eval = FALSE}
ALA_tags |>
write_csv(here("21mar2023reports.csv"))
```

# Summary example:

**Research Impact**

From March -- June 2023, the ALA has been referenced or mentioned in 134 published articles, with 91 of these articles directly using or citing hosted data in their analyses. All articles that cite the ALA are available on our website and are also now available on ALA Labs (including highlight summaries). The three main research foci for publications using the ALA this quarter were biosecurity, conservation, and citizen science.

Biosecurity was the largest focus, with ALA data used to illustrate invasion histories of several non-native species (McCulloch et al 2023). Introduction pathways, successful interventions, and overall pest impact on select species (such as the widespread forestry species Pinus radiata) are analysed at global scales (Brockerhoff et al., El Housni et al., Kachigunda et al., 2023; Ollivier et al., 2023). Steinrucken et al. (2023) also use ALA data to analyse the potential of fungal biocontrol agents against Sporobolus grasses in Australia, while, similarly, Yonow et al. (2023) use ALA data to model the potential of a parasite in reducing invasive snail populations.

Conservation research mostly analysed historical distributions, inferencing habitat preferences of currently at-risk species. Occurrence data contributed to constructions and reconstructions of the endangered Golden-shouldered Parrot in its range of the Cape York Peninsula (Crowley et al., 2023), as well as the greater bilby across eastern Australia (Silcock et al., 2023). Stock et al. (2023) find small inbreeding populations of the threatened orange-bellied tree frog; which are analysed further for their potential relationship to climate and environmental variables hosted by the Atlas.

Lastly, the capabilities of citizen science are emphasised in Djokic et al. (2023), who use DigiVol (hosted by the ALA) to allow volunteers to help identify species in images of historical fossils. Similarly, including analysis of image data in the ALA, Mesaglio et al. (2023) work to understand Australian regions with less photographed plant species' data and highlight the potential of citizen science to fill this gap for increased data richness and verifiability.

**Recent notable publications using the ALA:**

Brockerhoff, E. G., Gresham, B. A., Meurisse, N., Nahrung, H. F., Perret-Gentil, A., Pugh, A. R., Sopow, S. L., & Turner, R. M. (2023). Pining away and at home: Global utilisation of Pinus radiata by native and non-native insects. NeoBiota, 84, 137--167. https://doi.org/10.3897/neobiota.84.95864

Crowley, G. M., & Garnett, S. T. (2023). Distribution and decline of the Golden-shouldered Parrot Psephotellus chrysopterygius 1845--1990. North Queensland Naturalist, 53, 22--68.

Djokic, T., Frese, M., Woods, A., Dettmann, M., Flemons, P., Brink, F., & McCurry, M. R. (2023). Inferring the age and environmental characteristics of fossil sites using citizen science. PLOS ONE, 18(4), e0284388. https://doi.org/10.1371/journal.pone.0284388

EL Housni, Z., Abdessalem, T., Radouane, N., Ezrari, S., Zegoumou, A., & Ouijja, A. (2023). Overview of sugar beet leaf spot disease caused by Cercospora beticola Sacc. Archives of Phytopathology and Plant Protection, 56(7), 503--528. https://doi.org/10.1080/03235408.2023.2216356

Kachigunda, B., Coupland, G. T., Mengersen, K., Perera, D. I., van der Merwe, J., & McKirdy, S. (2023). A systems approach to biosecurity pays off: A case study investigating introduction pathways and interceptions of non-indigenous species at a biosecurity border. Biological Invasions. https://doi.org/10.1007/s10530-023-03077-5

McCulloch, G. A., Gurdasani, K., Hereward, J. P., Morin, L., Walter, G. H., & Raghu, S. (2023). Invasion history of Lycium ferocissimum in Australia: The impact of admixture on genetic diversity and differentiation. Diversity and Distributions, 1--13. https://doi.org/10.1111/ddi.13702

Mesaglio, T., Sauquet, H., Coleman, D., Wenk, E., & Cornwell, W. K. (2023). Photographs as an essential biodiversity resource: Drivers of gaps in the vascular plant photographic record. New Phytologist. https://doi.org/10.1111/nph.18813

Ollivier, M., Labouyrie, M., Raghu, S., Tavoillot, J., Tixier, M.-S., & Lesieur, V. (2023). Characterising the herbivore community and its impact on Sonchus oleraceus (Asterales: Asteraceae) in its invaded range in Australia. Austral Entomology, 1--15. https://doi.org/10.1111/aen.12640

Silcock, J. L., McRae, P. D., Laidlaw, M. J., & Southgate, R. I. (2023). Historical record shows broad habitat use and rapid decline of the greater bilby Macrotis lagotis in eastern Australia. Wildlife Research. https://doi.org/10.1071/WR22043

Steinrucken, T. V., & Vitelli, J. S. (2023). Biocontrol of weedy Sporobolus grasses in Australia using fungal pathogens. BioControl. https://doi.org/10.1007/s10526-023-10195-5

Stock, S. E., Klop-Toker, K., Wallace, S., Kelly, O., Callen, A., Seeto, R., Mahony, S. V., Hayward, M. W., & Mahony, M. J. (2023). Uncovering inbreeding, small populations, and strong genetic isolation in an Australian threatened frog, Litoria littlejohni. Conservation Genetics. https://doi.org/10.1007/s10592-023-01522-5

Yonow, T., Kriticos, D. J., Zalucki, M. P., Mc Donnell, R. J., & Caron, V. (2023). Population modelling for pest management: A case study using a pest land snail and its fly parasitoid in Australia. Ecological Modelling, 482, 110413. https://doi.org/10.1016/j.ecolmodel.2023.110413
