---
title: "Publication tracker reporting"
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
---

**Year in Review / Annual Report**

Once a year, the ALA produces a "Year in Review" report which encompasses our achievements over the last twelve months. This includes a section providing key summaries of the publication tracker, which differ from the quarterly report.\
\
They are:

1.  **The annual (calandar year) numbers of *journal articles* citing the ALA, up until the earliest article in 2007**
2.  **Annual (financial year) total publications (all) to journal article ratio**
3.  **Top 10 journal articles (grand total + yearly)**
4.  **Research domains or categories citing the ALA**

> This code was last updated on 2023-08-9.
>
> Data to run this code can be found [here](https://csiroau.sharepoint.com/:f:/r/sites/AtlasofLivingAustralia/Shared%20Documents/Science%20and%20Decision%20Support/Data/science/projects/literature-tracking/data?csf=1&web=1&e=M5hBEt), and within this repo: science - projects - literature-tracking - data (set to .gitignore).

# Step 1 - Data cleaning:

1.  Load required packages

```{r}
#| message: false
library(here)
library(tidyverse)
library(janitor)
library(lubridate)
```

2.  Load in exported .csv from Zotero

First you will need to export the latest encompassing .csv from Zotero and save it on your local system. You can do this by selecting **File \> Export library** in the desktop application.

```{r}
#| message: false
#| warning: false
data <- read_csv(here("projects", 
                      "literature-tracking", 
                      "data", 
                      "26 sept zotero 2023.csv"))
```

3.  Clean column names - it will make it easier to process and interpret the data.

```{r}
dataclean <- janitor ::clean_names(data) 
```

4.  We also want to make sure the date is the same across every entry so that it processes correctly when we filter it. `as_date` standardises all date data to `yyyy-mm-dd`. We also add a new column that extracts just the month and year for the month by month analysis below.

```{r}
dataclean2 <- dataclean |>
  mutate(
    date_added_clean = as_date(date_added),
    date_added_month = month(date_added_clean),
    date_added_year = year(date_added_clean)) |>
   select(title, date, place, publication_year, publication_title, 
          item_type, url, manual_tags, date_added, date_added_clean, 
          date_added_month, date_added_year)
dataclean2
```

# Step 2 - Summaries:

## Annual journal article numbers

In this section we collate the number of **journal articles** mentioning, citing, using, acknowledging etc. (...) the ALA in their given publication year.

> **Note that**: in the past we included the GBIF DOI tag in analyses.
>
> In 2022, it was decided to exclude this tag from the library (held on file) as GBIF's direct downloads are not always related to the ALA's own research impact.

```{r}
publicationbyyear <- dataclean2 |>
  filter(item_type == "journalArticle") |>
  group_by(publication_year) |>
  count() 
publicationbyyear
```

## Total to journal article ratio

*Total* publications within the financial year:

```{r}
totalpubsbyyear <- dataclean2 |>
  filter(date_added_clean >= ymd("2022-06-30")) |>
  filter(date_added_clean <= ymd("2023-06-30")) |>
  filter(publication_year == "2022" | publication_year == "2023") |>
  count() |>
  drop_na() 
totalpubsbyyear
```

No. of those publications that are *journal articles* within the financial year:

```{r}
journalratio <- dataclean2 |>
  filter(item_type == "journalArticle") |>
  filter(date_added_clean >= ymd("2022-06-30")) |>
  filter(date_added_clean <= ymd("2023-06-30")) |>
  filter(publication_year == "2022" | publication_year == "2023") |>
  count() |>
  drop_na() 
journalratio
```

## Top journal articles

This section analyses the ten most common journal articles mentioning, citing, using, acknowledging etc. the ALA **across all time**. It is worth checking that these journals are not preprint repositories (e.g. BioXriv), as, occasionally Zotero auto categorises them as journal articles instead.\
\
These numbers might change subtly over time due to ongoing data cleaning, but will likely remain similar. **They are included in the report**.

```{r}
grandtop10 <- dataclean2 |>
  filter(item_type == "journalArticle") |>
  group_by(publication_title) |>
  count() |>
   drop_na() |>
  arrange(desc(n)) 
grandtop10
```

```{r}
grandtop10$publication_title[1:10]
```

To find out how many different journal articles the ALA was cited in **the 2022-23 financial year**, run the same code but filter across the financial year dates via date_added (30 June - 30 June) and further restrict to papers published in that year (sometimes old papers are added as they become released online, which aren't relevant to our analysis).\
\
Be sure to remove the NA category (journal articles without an available name in the tracker) using `drop_na()` as well. This might be a chance to work on investigating these entries and see if there is missing data.

```{r}
top10currentyear2 <- dataclean2 |>
  filter(item_type == "journalArticle") |>
  filter(date_added_clean >= ymd("2022-06-30")) |>
  filter(date_added_clean <= ymd("2023-06-30")) |>
  filter(publication_year == "2022" | publication_year == "2023") |>
  group_by(publication_title) |>
  count() |>
  drop_na() |>
  arrange(desc(n)) 
top10currentyear2
```

**\
**Here, there are 189 rows: meaning 189 unique journal articles that have cited the ALA across this financial year. **This number is included in the report.**

The top journals for this financial year listed were:

```{r}
top10currentyear2$publication_title[1:10]
```

## Important filtering justifications

**We chose to filter by `date_added` and *then* restrict to `publication_year` instead of filtering by `date` (date of publication). Note that there are degrees of error in both methods for a few reasons:**

-   If filtered by `publication_year` and then restricted to the financial year dates `dd/mm/yyyy`, there is a risk of excluding papers that only have a *year* `yyyy` of publication listed (which is common). It is preferred to get the publication `date` of papers entered into the tracker closest to `dd/mm/yyyy` as possible.

-   Similarly, a paper may be published online but the listed publication `date` can be some months in advance as it reflects when the physical copy is released. While a paper may have been added and accessible online, it may be excluded from the financial year report if the physical release is outside of this date.

-   Using `date_added` and restricting to the relevant `publication_year`s, on the other hand, may exclude papers that were added outside of the financial year period but have a publication `date` within it. This seems to be the less likely occurrence and the more logical choice between the two.

### 2021 YiR error

1. GBIF papers

Note that for the 2021-2022 Annual Report/ Year in review (and those before) our numbers were retrieved with data that *still contained GBIF DOI tag papers* (on file as `30thjune21.csv`). Therefore, the number of publications in the '21-22 report are substantially larger across the entire dataset. 

GBIF DOI tagged papers are publications that have retrieved Australian biodiversity data from GBIF (our larger global node) but not the ALA. Technically this data is held in the ALA but as authors have retrieved it from the GBIF interface we decided that this is not a good metric of the ALA being 'used'.

You can find the GBIF papers removed from Zotero on file [here](https://csiroau.sharepoint.com/sites/AtlasofLivingAustralia/Shared%20Documents/Forms/AllItems.aspx?csf=1&web=1&e=M5hBEt&cid=7e5dbd4a%2D94f3%2D4356%2Dabe8%2D28c560fc0d3a&FolderCTID=0x01200060180DC552F0A64084CF983A43CEDF34&id=%2Fsites%2FAtlasofLivingAustralia%2FShared%20Documents%2FScience%20and%20Decision%20Support%2FData%2Fscience%2Fprojects%2Fliterature%2Dtracking%2Fdata%2Fremoved%20GBIF%20doi%20publications&viewid=669b2a1d%2Db1b9%2D4f7a%2D9a0f%2Db972843a2de4). 

2. Financial vs. calandar year

In addition to this, we *also* did not restrict the year to papers within the '21-'22 **financial year** specifically for the 21-22' report. We only presented **total to journal article ratios** in the full data **calendar year** of **2021**. The code below is what was used (using the `30thjune21.csv`). It is unclear whether this has always been the method of reporting during YiR as no documentation has been provided for these metrics. From '23 onwards we have decided to change this to the financial year due to it being a more recent metric. 

```{r}
#| code-fold: true
#| eval: false
dataclean2 |>
  select(publication_year, title, publication_title, item_type) |>
  filter(publication_year == 2021) |> 
  filter(item_type == "journalArticle") |>
  group_by(publication_title) |>
  count()
```

### Conclusion

Thinking about best practice for reporting in the data filtering points above, as well as financial year metrics being a more recent view of publications than the previous full available calendar year (as above, only '21 was provided for the '21-22 report as '21 at that point in time had a full year of publication data), we have made changes to how tracking data is filtered and reported. Additionally, having experienced issues replicating numbers from previous years we also decided to document all of our code for producing literature tracking metrics so that it can be easily reproduced and our justifications are clearly recorded for future decisions.


## Top research domains

We also include a research domain section that quantifies the the top 10 different fields of research the ALA has been utilized across. This involves text mining, and is a little more complex than the last analyses.\
\

To be added - Martin on file (?)
