---
title: "Australian Invasive Species"
author: "Atlas of Living Australia"
date: "`r Sys.Date()`"
always_allow_html: true
output:
  html_document:
    highlight: default
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
    include:
      in_header: template-files/header.html
      after_body: template-files/footer.html
    css: template-files/style.css
    self_contained: FALSE
editor_options: 
  chunk_output_type: console

---

```{r setup, include=FALSE}
library(htmltools)
source("template-files/functions.R")
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
knitr::opts_chunk$set(echo = FALSE)
```

------------------------------------------------------------------------------

This document uses data from the Atlas of Living Australia to provide counts and distributions of <br>
  - camels 
  - feral pigs 
  - feral goats
  - brumbies 
  - cane toads

------------------------------------------------------------------------------

```{r, message = FALSE, warning = FALSE}
# packages

library(tidyverse)
library(purrr)
library(janitor)
library(kableExtra)

# use ala_counts branch for multiple `group_by` and `select_filters` args
# remotes::install_github("AtlasOfLivingAustralia/galah@ala_counts")
library(galah)

# for mapping
library(ozmaps)
library(sf)
library(mapsf)

# Might need to install these:
# devtools::install_github('ropensci/rnaturalearthhires')
# install.packages("rgeos")
```

```{r, eval = FALSE}
# Don't forget to use galah_config:
# galah_config(email = "your_email@email.com")
```


# Camels {.tabset .tabset-fade}

## Yearly Observations

```{r, warning = FALSE, message = FALSE, echo = FALSE}
# select taxa and filters
taxa <- select_taxa("camelus")
filters <- select_filters(year > 2005)

# Download counts from ALA
total_camel_counts <- ala_counts(taxa = taxa)
camel_counts <- ala_counts(taxa = taxa, filters = filters, group_by = "year")

# Wrangle
camel_counts <- camel_counts %>% 
  mutate(
    year = as.integer(year)) %>%
  arrange(desc(year))
```


```{r, warning = FALSE, message = FALSE}
# Plot

# Fonts
library(extrafont)
library(showtext)
extrafont::loadfonts()
font_add_google(name = "Lato")

showtext_auto()

# Set font for {pilot} titles
pilot::set_pilot_family("Lato", title_family = "Lato") # Set font

# Counts by year
camel_yearly_counts <- camel_counts %>%
  ggplot(aes(x = year,
             y = count)) + 
  geom_bar(stat = "identity", fill = "#E06E53", width = 0.8) + 
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
                     expand = c(0,0)) +
  labs(x = "Year",
       y = "Number of Records",
       title = "Yearly Observations",
       subtitle = "(2006-2021)") +
  pilot::scale_fill_pilot() + 
  pilot::theme_pilot() + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 18),
        plot.title = element_text(size = 25, margin = margin(0,0,20,0)),
        plot.subtitle = element_text(size = 19, margin = margin(0,0,20,0)))

camel_yearly_counts

# pilot::add_pilot_titles(camel_yearly_counts, title = "Yearly Observations (2006-2021)")
```

```{r, warning = FALSE, message = FALSE, eval = FALSE}
# Table
camel_counts %>%
  rename_with(str_to_title) %>%
  kbl() %>%
  kable_styling("striped", full_width = FALSE) %>%
  kable_paper() %>%
  add_header_above(c("Yearly Counts" = 2))
```   

***


## Data Providers

```{r}
# Download data providers

camel_data_providers <- ala_counts(taxa = taxa, 
                             # filters = filters, 
                             group_by = "dataResourceName",
                             limit = 1e5)
```

```{r, echo = FALSE, results = FALSE}
# | Note: This table is set to not appear in knitted document
# | It is here in case exact numbers are requested, 
# | and can be added by removing `results = FALSE` in the chunk options

perc_camel_data_providers <- camel_data_providers %>%
  mutate(
    dataResourceName = str_trim(dataResourceName),
    percent = (count / sum(count)) * 100
  ) %>% 
  filter(percent > .01)

perc_camel_data_providers %>% 
  select(dataResourceName, count) %>%
  rename("Data Provider" = dataResourceName) %>%
  rename("Records" = count) %>%
  kbl() %>% kable_styling(full_width = FALSE)
```

```{r, fig.cap = "Note: Plot displays data providers that provide more than 0.1% of overall data in the ALA"}
# Plot

# Proportion of data from different data providers
plot_camel_data_providers <- perc_camel_data_providers %>% 
  filter(percent > .1) %>%
  
  ggplot(aes(x = reorder(dataResourceName, percent),
             y = percent,
             fill = dataResourceName,
             label = scales::percent(percent, suffix = "%"))) +
  geom_bar(stat = "identity") + 
  scale_x_discrete(labels = scales::label_wrap(35)) +
  scale_y_continuous(limits = c(0, 100), labels = function(x) paste0(x, "%")) + 
  pilot::scale_fill_pilot() + 
  pilot::theme_pilot(legend_position = "none") + 
  labs(x = "Data Provider",
       y = "Percentage of total records",
       title = "Data Providers") +
  coord_flip() + 
  theme(axis.text.y = element_text(size = 11),
        axis.text.x = element_text(size = 14),
        axis.title = element_text(size = 18),
        plot.title = element_text(size = 25, margin = margin(0,0,20,0)))

plot_camel_data_providers

# pilot::add_pilot_titles(plot_camel_data_providers, title = "Data Providers")
```

***

## Cartograph Map

```{r, eval = FALSE}
# Retrieve AUS map from {rnaturalearth}
aus_ne <- rnaturalearth::ne_states(country = "australia", returnclass = "sf")

# Counts of koalas by state
state_layers <- search_fields("australian states and territories") # id = "cl22"
camels_by_state <- ala_counts(taxa = taxa, group_by = "cl22")

# Rename column and trim "cl22" prefix
camels_by_state <- camels_by_state %>% 
  mutate(state = str_remove(cl22, "cl22:")) %>% 
  select(-cl22)

# join with data from {ozmaps}
camels_state_map <- merge(
  ozmaps::ozmap_states,
  camels_by_state,
  by.x = "NAME",
  by.y = "state"
)

# convert map coordinates to match ALA coordinates
camels_state_map <- st_transform(camels_state_map, crs = st_crs(3577))

par(mfrow=c(1,1))
# Map using mf_map() from {mapsf}

mf_theme(bg = "white")
mf_map(x = camels_state_map, var = "count", type = "choro",
       pal = "Dark Mint", border = "darkseagreen4", leg_pos = "n") %>%
  mf_label(var = "count", col = "#E06E53", cex = 2, overlap = FALSE, halo = TRUE, font = 2) <- mappy

```


## Density Map

```{r, message = FALSE, warning = FALSE, include = FALSE}
# Download occurrence records
camels_records <- ala_occurrences(taxa = taxa)

# Remove NAs
camels_records <- camels_records %>% filter(!is.na(.))

# filter records to only those on mainland/tasmania
camels_records_filtered <- camels_records %>%
  filter(decimalLongitude < 155,
         decimalLongitude > 110, 
         decimalLatitude > -45,
         decimalLatitude < -10)

# Plot
library(ggpointdensity)
showtext_auto()

camel_points_map <- ggplot(camels_records_filtered) + 
  geom_sf(data = ozmap_states, fill = "white") + 
  # geom_point(aes(x = decimalLongitude,
  #                y = decimalLatitude), color = "#B8573E") +
  geom_pointdensity(aes(x = decimalLongitude,
                        y = decimalLatitude)) +
  viridis::scale_color_viridis(option = "F", begin = 0.1) +
  # scale_colour_gradient(low = "#9E9E9F", high = "#B8573E") +
  labs(title = "Spatial Distribution",
       x = NULL, y = NULL) +
  pilot::theme_pilot(axes = "",
                     grid = "") + 
  guides(colour = guide_colourbar(title = "Number of \nOverlapping points"), alpha = "none") +
  theme(legend.position="bottom",
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,0,0),
        legend.title.align=0.5,
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        plot.title = element_text(size = 27, face = "bold", color = "#404040", hjust = 0.5, margin = margin(0,0,20,0)))
```


```{r, message = FALSE, warning = FALSE}
camel_points_map
```


## Hexagon map

Download data

```{r}
# Plot

camel_hex_map <- ggplot(camels_records_filtered) +
  geom_sf(data = ozmap_states, fill = "white") +
  geom_hex(mapping = aes(x = decimalLongitude, 
                         y = decimalLatitude,
                         alpha=log(..count..)), bins = 50) +
  ylim(-45, -10) +
  xlim(110, 155) +
  # viridis::scale_fill_viridis(option = "E") +
  scale_fill_gradientn(colours = c("#FFC557", "#B8573E")) +
  theme_bw() +
  theme(
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    legend.position = "bottom") + 
  guides(fill = guide_colourbar(title = "Count"), alpha = "none")

camel_hex_map
```

## Final Plot

```{r}
library(patchwork)

camels <- ((camel_yearly_counts / plot_camel_data_providers) | camel_points_map) + 
  plot_layout(widths = c(1, 4), heights = c(1, 4))

camel_title <- expression(paste("Camels (", italic("Camelus"), ")", sep = ""))
camel_subtitle <- paste("Atlas of Living Australia (Total number of records = ", total_camel_counts, ")")

camels <- pilot::add_pilot_titles(camels,
                                  title = camel_title,
                                  subtitle = camel_subtitle,
                                  title_family = "Lato",
                                  title_size = 30,
                                  title_color = element_text(color = "#E06E53"),
                                  subtitle_color = "#B8573E")
camels
```

```{r, eval = FALSE}
ggsave(here::here("projects", "invasive-species-counts", "plots", "camels.pdf"),
       dpi = 600, width = 20, height = 13, units = "in")
```


# Feral Pigs

## Yearly Observations

```{r, warning = FALSE, message = FALSE, echo = FALSE}
# select taxa and filters
taxa <- select_taxa("sus scrofa")
filters <- select_filters(year > 2005)

# Download counts from ALA
total_pig_counts <- ala_counts(taxa = taxa)

pig_counts <- ala_counts(taxa = taxa, filters = filters, group_by = "year")

# Wrangle
pig_counts <- pig_counts %>% 
  mutate(
    year = as.integer(year)) %>%
  arrange(desc(year))
```


```{r, warning = FALSE, message = FALSE}
# Plot

showtext_auto()

# Set font for {pilot} titles
pilot::set_pilot_family("Lato", title_family = "Lato") # Set font

# Counts by year
pig_yearly_counts <- pig_counts %>%
  ggplot(aes(x = year,
             y = count)) + 
  geom_bar(stat = "identity", fill = "#E06E53", width = 0.8) + 
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
                     expand = c(0,0)) +
  labs(x = "Year",
       y = "Number of Records",
       title = "Yearly Observations",
       subtitle = "(2006-2021)") +
  pilot::scale_fill_pilot() + 
  pilot::theme_pilot() + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 18),
        plot.title = element_text(size = 25, margin = margin(0,0,20,0)),
        plot.subtitle = element_text(size = 19, margin = margin(0,0,20,0)))

pig_yearly_counts

# pilot::add_pilot_titles(camel_yearly_counts, title = "Yearly Observations (2006-2021)")
```

```{r, warning = FALSE, message = FALSE, eval = FALSE}
# Table
pig_counts %>%
  rename_with(str_to_title) %>%
  kbl() %>%
  kable_styling("striped", full_width = FALSE) %>%
  kable_paper() %>%
  add_header_above(c("Yearly Counts" = 2))
```   

***


## Data Providers

```{r}
# Download data providers

pig_data_providers <- ala_counts(taxa = taxa, 
                             # filters = filters, 
                             group_by = "dataResourceName",
                             limit = 1e5)
```

```{r, echo = FALSE, results = FALSE}
# | Note: This table is set to not appear in knitted document
# | It is here in case exact numbers are requested, 
# | and can be added by removing `results = FALSE` in the chunk options

perc_pig_data_providers <- pig_data_providers %>%
  mutate(
    dataResourceName = str_trim(dataResourceName),
    percent = (count / sum(count)) * 100
  ) %>% 
  filter(percent > .01)

perc_pig_data_providers %>% 
  select(dataResourceName, count) %>%
  rename("Data Provider" = dataResourceName) %>%
  rename("Records" = count) %>%
  kbl() %>% kable_styling(full_width = FALSE)
```

```{r, fig.cap = "Note: Plot displays data providers that provide more than 0.1% of overall data in the ALA"}
# Plot

# Proportion of data from different data providers
plot_pig_data_providers <- perc_pig_data_providers %>% 
  filter(percent > 0.5) %>%
  
  ggplot(aes(x = reorder(dataResourceName, percent),
             y = percent,
             fill = dataResourceName,
             label = scales::percent(percent, suffix = "%"))) +
  geom_bar(stat = "identity") + 
  scale_x_discrete(labels = scales::label_wrap(35)) +
  scale_y_continuous(limits = c(0, 100), labels = function(x) paste0(x, "%")) + 
  pilot::scale_fill_pilot() + 
  pilot::theme_pilot(legend_position = "none") + 
  labs(x = "Data Provider",
       y = "Percentage of total records",
       title = "Data Providers") +
  coord_flip() + 
  theme(axis.text.y = element_text(size = 10),
        axis.text.x = element_text(size = 14),
        axis.title = element_text(size = 18),
        plot.title = element_text(size = 25, margin = margin(0,0,20,0)))

plot_pig_data_providers

# pilot::add_pilot_titles(plot_camel_data_providers, title = "Data Providers")
```

***


## Density Map

```{r, message = FALSE, warning = FALSE, include = FALSE}
# Download occurrence records
pig_records <- ala_occurrences(taxa = taxa)

# Remove NAs
pig_records <- pig_records %>% filter(!is.na(.))

# filter records to only those on mainland/tasmania
pig_records_filtered <- pig_records %>%
  filter(decimalLongitude < 155,
         decimalLongitude > 110, 
         decimalLatitude > -45,
         decimalLatitude < -10)

# Plot
library(ggpointdensity)
showtext_auto()

pig_points_map <- ggplot(pig_records_filtered) + 
  geom_sf(data = ozmap_states, fill = "white") + 
  # geom_point(aes(x = decimalLongitude,
  #                y = decimalLatitude), color = "#B8573E") +
  geom_pointdensity(aes(x = decimalLongitude,
                        y = decimalLatitude)) +
  viridis::scale_color_viridis(option = "F", 
                               begin = 0.1, 
                               breaks = c(1000, 3000, 5000)) +
  # scale_colour_gradient(low = "#9E9E9F", high = "#B8573E") +
  labs(title = "Spatial Distribution",
       x = NULL, y = NULL) +
  pilot::theme_pilot(axes = "",
                     grid = "") + 
  guides(colour = guide_colourbar(title = "Number of \nOverlapping points"), alpha = "none") +
  theme(legend.position="bottom",
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,0,0),
        legend.title.align=0.5,
        legend.text = element_text(angle = 0),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        plot.title = element_text(size = 27, face = "bold", color = "#404040", hjust = 0.5, margin = margin(0,0,20,0)))
```


```{r, message = FALSE, warning = FALSE}
pig_points_map
```

***

## Final Plot

```{r}
library(patchwork)

pigs <- ((pig_yearly_counts / plot_pig_data_providers) | pig_points_map) + 
  plot_layout(widths = c(1, 4), heights = c(1, 4))

pig_title <- expression(paste("Feral Pigs (", italic("Sus scrofa"), ")", sep = ""))
pig_subtitle <- paste("Atlas of Living Australia (Total number of records = ", total_pig_counts, ")")

feral_pigs <- pilot::add_pilot_titles(pigs,
                                  title = pig_title,
                                  subtitle = pig_subtitle,
                                  title_family = "Lato",
                                  title_size = 30,
                                  title_color = element_text(color = "#E06E53"),
                                  subtitle_color = "#B8573E")
feral_pigs
```

```{r, eval = FALSE}
ggsave(here::here("projects", "invasive-species-counts", "plots", "feral-pigs.pdf"),
       dpi = 600, width = 20, height = 13, units = "in")
```







# Feral Goats

## Yearly Observations

```{r, warning = FALSE, message = FALSE, echo = FALSE}
# select taxa and filters
taxa <- select_taxa("capra hircus")
filters <- select_filters(year > 2005)

# Download counts from ALA
total_goat_counts <- ala_counts(taxa = taxa)

goat_counts <- ala_counts(taxa = taxa, filters = filters, group_by = "year")

# Wrangle
goat_counts <- goat_counts %>% 
  mutate(
    year = as.integer(year)) %>%
  arrange(desc(year))
```


```{r, warning = FALSE, message = FALSE}
# Plot

showtext_auto()

# Set font for {pilot} titles
pilot::set_pilot_family("Lato", title_family = "Lato") # Set font

# Counts by year
goat_yearly_counts <- goat_counts %>%
  ggplot(aes(x = year,
             y = count)) + 
  geom_bar(stat = "identity", fill = "#E06E53", width = 0.8) + 
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
                     expand = c(0,0)) +
  labs(x = "Year",
       y = "Number of Records",
       title = "Yearly Observations",
       subtitle = "(2006-2021)") +
  pilot::scale_fill_pilot() + 
  pilot::theme_pilot() + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 18),
        plot.title = element_text(size = 25, margin = margin(0,0,20,0)),
        plot.subtitle = element_text(size = 19, margin = margin(0,0,20,0)))

goat_yearly_counts

# pilot::add_pilot_titles(camel_yearly_counts, title = "Yearly Observations (2006-2021)")
```

```{r, warning = FALSE, message = FALSE, eval = FALSE}
# Table
goat_counts %>%
  rename_with(str_to_title) %>%
  kbl() %>%
  kable_styling("striped", full_width = FALSE) %>%
  kable_paper() %>%
  add_header_above(c("Yearly Counts" = 2))
```   

***


## Data Providers

```{r}
# Download data providers

goat_data_providers <- ala_counts(taxa = taxa, 
                             # filters = filters, 
                             group_by = "dataResourceName",
                             limit = 1e5)
```

```{r, echo = FALSE, results = FALSE}
# | Note: This table is set to not appear in knitted document
# | It is here in case exact numbers are requested, 
# | and can be added by removing `results = FALSE` in the chunk options

perc_goat_data_providers <- goat_data_providers %>%
  mutate(
    dataResourceName = str_trim(dataResourceName),
    percent = (count / sum(count)) * 100
  ) %>% 
  filter(percent > .01)

perc_goat_data_providers %>% 
  select(dataResourceName, count) %>%
  rename("Data Provider" = dataResourceName) %>%
  rename("Records" = count) %>%
  kbl() %>% kable_styling(full_width = FALSE)
```

```{r, fig.cap = "Note: Plot displays data providers that provide more than 0.1% of overall data in the ALA"}
# Plot

# Proportion of data from different data providers
plot_goat_data_providers <- perc_goat_data_providers %>% 
  filter(percent > 0.1) %>%
  
  ggplot(aes(x = reorder(dataResourceName, percent),
             y = percent,
             fill = dataResourceName,
             label = scales::percent(percent, suffix = "%"))) +
  geom_bar(stat = "identity") + 
  scale_x_discrete(labels = scales::label_wrap(35)) +
  scale_y_continuous(limits = c(0, 100), labels = function(x) paste0(x, "%")) + 
  pilot::scale_fill_pilot() + 
  pilot::theme_pilot(legend_position = "none") + 
  labs(x = "Data Provider",
       y = "Percentage of total records",
       title = "Data Providers") +
  coord_flip() + 
  theme(axis.text.y = element_text(size = 10),
        axis.text.x = element_text(size = 14),
        axis.title = element_text(size = 18),
        plot.title = element_text(size = 25, margin = margin(0,0,20,0)))

plot_goat_data_providers

# pilot::add_pilot_titles(plot_camel_data_providers, title = "Data Providers")
```

***


## Density Map

```{r, message = FALSE, warning = FALSE, include = FALSE}
# Download occurrence records
goat_records <- ala_occurrences(taxa = taxa)

# Remove NAs
goat_records <- goat_records %>% filter(!is.na(.))

# filter records to only those on mainland/tasmania
goat_records_filtered <- goat_records %>%
  filter(decimalLongitude < 155,
         decimalLongitude > 110, 
         decimalLatitude > -45,
         decimalLatitude < -10)

# Plot
library(ggpointdensity)
showtext_auto()

goat_points_map <- ggplot(goat_records_filtered) + 
  geom_sf(data = ozmap_states, fill = "white") + 
  # geom_point(aes(x = decimalLongitude,
  #                y = decimalLatitude), color = "#B8573E") +
  geom_pointdensity(aes(x = decimalLongitude,
                        y = decimalLatitude)) +
  viridis::scale_color_viridis(option = "F", 
                               begin = 0.1, 
                               breaks = c(1000, 4000, 7000)) +
  # scale_colour_gradient(low = "#9E9E9F", high = "#B8573E") +
  labs(title = "Spatial Distribution",
       x = NULL, y = NULL) +
  pilot::theme_pilot(axes = "",
                     grid = "") + 
  guides(colour = guide_colourbar(title = "Number of \nOverlapping points"), alpha = "none") +
  theme(legend.position="bottom",
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,0,0),
        legend.title.align=0.5,
        legend.text = element_text(angle = 0),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        plot.title = element_text(size = 27, face = "bold", color = "#404040", hjust = 0.5, margin = margin(0,0,20,0)))
```


```{r, message = FALSE, warning = FALSE}
goat_points_map
```

***

## Final Plot

```{r}
library(patchwork)

goats <- ((goat_yearly_counts / plot_goat_data_providers) | goat_points_map) + 
  plot_layout(widths = c(1, 4), heights = c(1, 4))

goat_title <- expression(paste("Feral Goats (", italic("Capra hircus"), ")", sep = ""))
goat_subtitle <- paste("Atlas of Living Australia (Total number of records = ", total_goat_counts, ")")

feral_goats <- pilot::add_pilot_titles(goats,
                                  title = goat_title,
                                  subtitle = goat_subtitle,
                                  title_family = "Lato",
                                  title_size = 30,
                                  title_color = element_text(color = "#E06E53"),
                                  subtitle_color = "#B8573E")
feral_goats
```

```{r, eval = FALSE}
ggsave(here::here("projects", "invasive-species-counts", "plots", "feral-goats.pdf"),
       dpi = 600, width = 20, height = 13, units = "in")
```






# Brumbies

## Yearly Observations

```{r, warning = FALSE, message = FALSE, echo = FALSE}
# select taxa and filters
taxa <- select_taxa("equus caballus")
filters <- select_filters(year > 2005)

# Download counts from ALA
total_brumby_counts <- ala_counts(taxa = taxa)

brumby_counts <- ala_counts(taxa = taxa, filters = filters, group_by = "year")

# Wrangle
brumby_counts <- brumby_counts %>% 
  mutate(
    year = as.integer(year)) %>%
  arrange(desc(year))
```


```{r, warning = FALSE, message = FALSE}
# Plot

showtext_auto()

# Set font for {pilot} titles
pilot::set_pilot_family("Lato", title_family = "Lato") # Set font

# Counts by year
brumby_yearly_counts <- brumby_counts %>%
  ggplot(aes(x = year,
             y = count)) + 
  geom_bar(stat = "identity", fill = "#E06E53", width = 0.8) + 
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
                     expand = c(0,0)) +
  labs(x = "Year",
       y = "Number of Records",
       title = "Yearly Observations",
       subtitle = "(2006-2021)") +
  pilot::scale_fill_pilot() + 
  pilot::theme_pilot() + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 18),
        plot.title = element_text(size = 25, margin = margin(0,0,20,0)),
        plot.subtitle = element_text(size = 19, margin = margin(0,0,20,0)))

brumby_yearly_counts

# pilot::add_pilot_titles(camel_yearly_counts, title = "Yearly Observations (2006-2021)")
```

```{r, warning = FALSE, message = FALSE, eval = FALSE}
# Table
brumby_counts %>%
  rename_with(str_to_title) %>%
  kbl() %>%
  kable_styling("striped", full_width = FALSE) %>%
  kable_paper() %>%
  add_header_above(c("Yearly Counts" = 2))
```   

***


## Data Providers

```{r}
# Download data providers

brumby_data_providers <- ala_counts(taxa = taxa, 
                             # filters = filters, 
                             group_by = "dataResourceName",
                             limit = 1e5)
```

```{r, echo = FALSE, results = FALSE}
# | Note: This table is set to not appear in knitted document
# | It is here in case exact numbers are requested, 
# | and can be added by removing `results = FALSE` in the chunk options

perc_brumby_data_providers <- brumby_data_providers %>%
  mutate(
    dataResourceName = str_trim(dataResourceName),
    percent = (count / sum(count)) * 100
  ) %>% 
  filter(percent > .01)

perc_brumby_data_providers %>% 
  select(dataResourceName, count) %>%
  rename("Data Provider" = dataResourceName) %>%
  rename("Records" = count) %>%
  kbl() %>% kable_styling(full_width = FALSE)
```

```{r, fig.cap = "Note: Plot displays data providers that provide more than 0.1% of overall data in the ALA"}
# Plot

# Proportion of data from different data providers
plot_brumby_data_providers <- perc_brumby_data_providers %>% 
  filter(percent > 0.1) %>%
  
  ggplot(aes(x = reorder(dataResourceName, percent),
             y = percent,
             fill = dataResourceName,
             label = scales::percent(percent, suffix = "%"))) +
  geom_bar(stat = "identity") + 
  scale_x_discrete(labels = scales::label_wrap(35)) +
  scale_y_continuous(limits = c(0, 100), labels = function(x) paste0(x, "%")) + 
  pilot::scale_fill_pilot() + 
  pilot::theme_pilot(legend_position = "none") + 
  labs(x = "Data Provider",
       y = "Percentage of total records",
       title = "Data Providers") +
  coord_flip() + 
  theme(axis.text.y = element_text(size = 10),
        axis.text.x = element_text(size = 14),
        axis.title = element_text(size = 18),
        plot.title = element_text(size = 25, margin = margin(0,0,20,0)))

plot_brumby_data_providers

# pilot::add_pilot_titles(plot_camel_data_providers, title = "Data Providers")
```

***


## Density Map

```{r, message = FALSE, warning = FALSE, include = FALSE}
# Download occurrence records
brumby_records <- ala_occurrences(taxa = taxa)

# Remove NAs
brumby_records <- brumby_records %>% filter(!is.na(.))

# filter records to only those on mainland/tasmania
brumby_records_filtered <- brumby_records %>%
  filter(decimalLongitude < 155,
         decimalLongitude > 110, 
         decimalLatitude > -45,
         decimalLatitude < -10)

# Plot
library(ggpointdensity)
showtext_auto()

brumby_points_map <- ggplot(brumby_records_filtered) + 
  geom_sf(data = ozmap_states, fill = "white") + 
  # geom_point(aes(x = decimalLongitude,
  #                y = decimalLatitude), color = "#B8573E") +
  geom_pointdensity(aes(x = decimalLongitude,
                        y = decimalLatitude)) +
  viridis::scale_color_viridis(option = "F", 
                               begin = 0.1, 
                               breaks = c(100, 1000, 2000)) +
  # scale_colour_gradient(low = "#9E9E9F", high = "#B8573E") +
  labs(title = "Spatial Distribution",
       x = NULL, y = NULL) +
  pilot::theme_pilot(axes = "",
                     grid = "") + 
  guides(colour = guide_colourbar(title = "Number of \nOverlapping points"), alpha = "none") +
  theme(legend.position="bottom",
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,0,0),
        legend.title.align=0.5,
        legend.text = element_text(angle = 0),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        plot.title = element_text(size = 27, face = "bold", color = "#404040", hjust = 0.5, margin = margin(0,0,20,0)))
```


```{r, message = FALSE, warning = FALSE}
brumby_points_map
```

***


## Final Plot

```{r}
library(patchwork)

brumbys <- ((brumby_yearly_counts / plot_brumby_data_providers) | brumby_points_map) + 
  plot_layout(widths = c(1, 4), heights = c(1, 4))

brumby_title <- expression(paste("Feral Brumbies (", italic("Equus caballus"), ")", sep = ""))
brumby_subtitle <- paste("Atlas of Living Australia (Total number of records = ", total_brumby_counts, ")")

feral_brumbys <- pilot::add_pilot_titles(brumbys,
                                  title = brumby_title,
                                  subtitle = brumby_subtitle,
                                  title_family = "Lato",
                                  title_size = 30,
                                  title_color = element_text(color = "#E06E53"),
                                  subtitle_color = "#B8573E")
feral_brumbys
```

```{r, eval = FALSE}
ggsave(here::here("projects", "invasive-species-counts", "plots", "brumbies.pdf"),
       dpi = 600, width = 20, height = 13, units = "in")
```





# Cane Toads

## Yearly Observations

```{r, warning = FALSE, message = FALSE, echo = FALSE}
# select taxa and filters
taxa <- select_taxa("Rhinella marina")
filters <- select_filters(year > 2005)

# Download counts from ALA
total_toad_counts <- ala_counts(taxa = taxa)

toad_counts <- ala_counts(taxa = taxa, filters = filters, group_by = "year")

# Wrangle
toad_counts <- toad_counts %>% 
  mutate(
    year = as.integer(year)) %>%
  arrange(desc(year))
```


```{r, warning = FALSE, message = FALSE}
# Plot

showtext_auto()

# Set font for {pilot} titles
pilot::set_pilot_family("Lato", title_family = "Lato") # Set font

# Counts by year
toad_yearly_counts <- toad_counts %>%
  ggplot(aes(x = year,
             y = count)) + 
  geom_bar(stat = "identity", fill = "#E06E53", width = 0.8) + 
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
                     expand = c(0,0)) +
  labs(x = "Year",
       y = "Number of Records",
       title = "Yearly Observations",
       subtitle = "(2006-2021)") +
  pilot::scale_fill_pilot() + 
  pilot::theme_pilot() + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 18),
        plot.title = element_text(size = 25, margin = margin(0,0,20,0)),
        plot.subtitle = element_text(size = 19, margin = margin(0,0,20,0)))

toad_yearly_counts

# pilot::add_pilot_titles(camel_yearly_counts, title = "Yearly Observations (2006-2021)")
```

```{r, warning = FALSE, message = FALSE, eval = FALSE}
# Table
toad_counts %>%
  rename_with(str_to_title) %>%
  kbl() %>%
  kable_styling("striped", full_width = FALSE) %>%
  kable_paper() %>%
  add_header_above(c("Yearly Counts" = 2))
```   

***


## Data Providers

```{r}
# Download data providers

toad_data_providers <- ala_counts(taxa = taxa, 
                             # filters = filters, 
                             group_by = "dataResourceName",
                             limit = 1e5)
```

```{r, echo = FALSE, results = FALSE}
# | Note: This table is set to not appear in knitted document
# | It is here in case exact numbers are requested, 
# | and can be added by removing `results = FALSE` in the chunk options

perc_toad_data_providers <- toad_data_providers %>%
  mutate(
    dataResourceName = str_trim(dataResourceName),
    percent = (count / sum(count)) * 100
  ) %>% 
  filter(percent > .01)

perc_toad_data_providers %>% 
  select(dataResourceName, count) %>%
  rename("Data Provider" = dataResourceName) %>%
  rename("Records" = count) %>%
  kbl() %>% kable_styling(full_width = FALSE)
```

```{r, fig.cap = "Note: Plot displays data providers that provide more than 0.1% of overall data in the ALA"}
# Plot

# Proportion of data from different data providers
plot_toad_data_providers <- perc_toad_data_providers %>% 
  filter(percent > 1) %>%
  
  ggplot(aes(x = reorder(dataResourceName, percent),
             y = percent,
             fill = dataResourceName,
             label = scales::percent(percent, suffix = "%"))) +
  geom_bar(stat = "identity") + 
  scale_x_discrete(labels = scales::label_wrap(35)) +
  scale_y_continuous(limits = c(0, 100), labels = function(x) paste0(x, "%")) + 
  pilot::scale_fill_pilot() + 
  pilot::theme_pilot(legend_position = "none") + 
  labs(x = "Data Provider",
       y = "Percentage of total records",
       title = "Data Providers") +
  coord_flip() + 
  theme(axis.text.y = element_text(size = 10),
        axis.text.x = element_text(size = 14),
        axis.title = element_text(size = 18),
        plot.title = element_text(size = 25, margin = margin(0,0,20,0)))

plot_toad_data_providers

# pilot::add_pilot_titles(plot_camel_data_providers, title = "Data Providers")
```

***


## Density Map

```{r, message = FALSE, warning = FALSE, include = FALSE}
# Download occurrence records
toad_records <- ala_occurrences(taxa = taxa)

# Remove NAs
toad_records <- toad_records %>% filter(!is.na(.))

# filter records to only those on mainland/tasmania
toad_records_filtered <- toad_records %>%
  filter(decimalLongitude < 155,
         decimalLongitude > 110, 
         decimalLatitude > -45,
         decimalLatitude < -10)

# Plot
library(ggpointdensity)
showtext_auto()

toad_points_map <- ggplot(toad_records_filtered) + 
  geom_sf(data = ozmap_states, fill = "white") + 
  # geom_point(aes(x = decimalLongitude,
  #                y = decimalLatitude), color = "#B8573E") +
  geom_pointdensity(aes(x = decimalLongitude,
                        y = decimalLatitude)) +
  viridis::scale_color_viridis(option = "F", 
                               begin = 0.1, 
                               breaks = c(1000, 2500, 4000)) +
  # scale_colour_gradient(low = "#9E9E9F", high = "#B8573E") +
  labs(title = "Spatial Distribution",
       x = NULL, y = NULL) +
  pilot::theme_pilot(axes = "",
                     grid = "") + 
  guides(colour = guide_colourbar(title = "Number of \nOverlapping points"), alpha = "none") +
  theme(legend.position="bottom",
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,0,0),
        legend.title.align=0.5,
        legend.text = element_text(angle = 0),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        plot.title = element_text(size = 27, face = "bold", color = "#404040", hjust = 0.5, margin = margin(0,0,20,0)))
```


```{r, message = FALSE, warning = FALSE}
toad_points_map
```

***


## Final Plot

```{r}
library(patchwork)

toads <- ((toad_yearly_counts / plot_toad_data_providers) | toad_points_map) + 
  plot_layout(widths = c(1, 4), heights = c(1, 4))

toad_title <- expression(paste("Cane Toads (", italic("Rhinella marina"), ")", sep = ""))
toad_subtitle <- paste("Atlas of Living Australia (Total number of records = ", total_toad_counts, ")")

feral_toads <- pilot::add_pilot_titles(toads,
                                  title = toad_title,
                                  subtitle = toad_subtitle,
                                  title_family = "Lato",
                                  title_size = 30,
                                  title_color = element_text(color = "#E06E53"),
                                  subtitle_color = "#B8573E")
feral_toads
```

```{r, eval = FALSE}
ggsave(here::here("projects", "invasive-species-counts", "plots", "cane-toads.pdf"),
       dpi = 600, width = 20, height = 13, units = "in")
```
